---
title: "Crime Exploratory Data Analysis"
author: "Erin Franke"
date: "2022-10-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Loading 

```{r}
# Libraries 
library(tidyverse)
library(NatParksPalettes)
library(PrettyCols)
library(ggmap)
library(sf)

load('Data/tractData/cleaned/gent_clean.RData')

#Minneapolis
load('Data/CrimeCleaned/tract_aggregated_minneapolis.RData')
load('Data/CrimeCleaned/tract_points_minneapolis.RData')
load("Data/CrimeCleaned/tract_aggregated_minneapolis_violent.RData")
load("Data/CrimeCleaned/tract_aggregated_minneapolis_theft.RData")
load("Data/CrimeCleaned/tract_aggregated_minneapolis_autotheft.RData")


#Saint Paul
load('Data/CrimeCleaned/tract_aggregated_sp.RData')
load("Data/CrimeCleaned/tract_points_sp.RData")
load("Data/CrimeCleaned/tract_aggregated_sp_violent.RData")
load("Data/CrimeCleaned/tract_aggregated_sp_theft.RData")
load("Data/CrimeCleaned/tract_aggregated_sp_autotheft.RData")
```

# EDA - Minneapolis

Explore year distribution of crimes in Minneapolis.

```{r}
mn2019 <- tract_aggregated %>%
  filter(year == 2019) %>%
  select(crimes_per_1000d)
sp2019 <- tract_aggregated_sp %>%
  filter(year == 2019) %>%
  select(crimes_per_1000d)

bind_rows(mn2019, sp2019) %>%
  ggplot()+
  geom_sf(aes(fill = crimes_per_1000d))+
  theme_classic()+
  scale_fill_natparks_d(name = "Olympic")+
  labs(title = "Crimes per 1000 residents in 2019 by census tract", fill = "", y = "", x = "")+
  annotate("point", x = -93.1691, y = 44.938, colour = "white", size = 1) +
  annotate(geom = "text", x = -93.1691, y = 44.945, label = "Macalester \nCollege", size = 2, color = "white")+
  annotate("point", x = -93.1022, y = 44.955, colour = "white", size = 1) +
  annotate(geom = "text", x = -93.1022, y = 44.9525, label = "Capitol", size = 2, color = "white")+
  annotate("point", x = -93.2776, y = 44.9817, colour = "black", size = 1) +
  annotate(geom = "text", x = -93.285, y = 44.9817, label = "Target \nField", size = 2.5, color = "black")+
  annotate("point", x = -93.2732, y = 44.9585, colour = "black", size = 1) +
  annotate(geom = "text", x = -93.2732, y = 44.9615, label = "MIA", size = 2.5, color = "black")+
  theme(plot.title.position = "plot", 
        plot.title = element_text(family = "mono", size = 12), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(), 
        axis.line = element_blank(), 
        legend.text = element_text(family = "mono"))
```

Look at overall trend in number of crimes per year.

```{r}
minn_annual <- tract_points %>%
  st_drop_geometry() %>%
  count(year) %>%
  mutate(city = "Minneapolis") %>%
  filter(year < 2022)

sp_annual <- tract_points_sp %>%
  st_drop_geometry() %>%
  count(year) %>%
  mutate(city = "Saint Paul")

bind_rows(minn_annual, sp_annual) %>%
  ggplot(aes(x=year, y=n, color = city))+
  geom_point()+
  geom_line()+
  ylim(c(5000, 25000))+
    scale_color_manual(values = c("navy", "goldenrod3"))+
  scale_x_continuous(breaks = c(2010, 2012, 2014, 2016, 2018, 2020, 2022))+
  theme_classic()+
  labs(x="Year", y= "Number of crimes", title = "Annual Crimes in the Twin Cities", color = "City")+
  theme(plot.title.position = "plot", 
        plot.title = element_text(family = "mono", size =12), 
        axis.title = element_text(family = "mono"), 
        legend.title = element_text(family = "mono"), 
        legend.text = element_text(family = "mono"), 
        axis.text = element_text(family = "mono"))
```

```{r}
tract_aggregated %>%
  st_drop_geometry() %>%
  group_by(year, canGent) %>%
  summarize(avg = mean(crimes_per_1000)) %>%
  mutate(year = as.numeric(year)) %>%
  filter(year < 2022) %>%
  mutate(canGent = case_when(canGent == "1" ~ "Gentrifiable", 
                             TRUE ~ "Cannot gentrify")) %>%
  ggplot(aes(x=year, y=avg, color = as.factor(canGent)))+
  geom_point()+
  geom_line()+
    scale_color_manual(values = c("navy", "goldenrod3"))+
  scale_x_continuous(breaks = c(2010, 2012, 2014, 2016, 2018, 2020, 2022))+
  theme_classic()+
  labs(x="Year", y="Crimes per 1000 residents" , title = "Annual crimes per 1000 residents in Minneapolis by", subtitle = "2010 tract gentrification status", color = "")+
  theme(plot.title.position = "plot", 
        plot.title = element_text(family = "mono", size =12),
        plot.subtitle = element_text(family = "mono", size =12), 
        axis.title = element_text(family = "mono"), 
        legend.title = element_text(family = "mono"), 
        legend.text = element_text(family = "mono"), 
        axis.text = element_text(family = "mono"))
```



Spaghetti plot

```{r}
sp_annual_tract <- tract_points_sp %>%
  st_drop_geometry() %>%
  count(year, tractmod) %>%
  filter(year != 2022, year != 2014) %>%
  mutate(City = "Saint Paul") %>%
  rename("tract" = "tractmod")

minn_annual_tract <- tract_points %>%
  st_drop_geometry() %>%
  count(year, tract) %>%
  filter(year != 2022) %>%
  mutate(City = "Minneapolis")

bind_rows(minn_annual_tract, sp_annual_tract) %>%
  filter(tract != 1044, tract !=1261) %>%
  ggplot(aes(x=year, y=n, group = tract))+
  geom_line(color = "gray")+
  scale_x_continuous(breaks = c(2010, 2012, 2014, 2016, 2018, 2020, 2022))+
  facet_wrap(~City, scales = "free_x")+
  theme_classic()+
  labs(x="Year", y= "Recorded crimes", title = "Annual recorded crimes by census tract")+
  theme(plot.title.position = "plot", 
        plot.title = element_text(family = "mono", size =12),
        axis.text = element_text(family = "mono"),
        axis.title = element_text(family = "mono"), 
        legend.title = element_text(family = "mono"), 
        legend.text = element_text(family = "mono"), 
        strip.background = element_blank(), 
        strip.text = element_text(family = "mono", size = 10))
```


Create groups of offenses and plot trends over years. \
- Theft: THEFT, BIKETF, TFPER, THFTSW, PKGTHEFT \
- Autotheft: TFMV, AUTOTH, TMVP, CARJCK, MVTHFT \
- Home burglary: BURGD \
- Business burglary: BURGB, ROBBIZ, TBLDG \
- Assault: ASLT2, CSCR, DASTR, DASLT2, ASLT3, ASLT4, DASLT3, MURDR,ASLT1, DASLT1 \

```{r}
tract_points %>%
  st_drop_geometry() %>%
  count(year, offense_group) %>%
  filter(offense_group != "Other", year < 2022) %>%
  ggplot(aes(x=year, y=n, color = offense_group))+
  geom_point()+
  geom_line()+
  theme_classic()+
  labs(x="Year", y= "", title = "Number of Recorded Minneapolis Crimes by Year", color = "Crime category")+
  theme(plot.title.position = "plot", 
        plot.title = element_text(family = "mono", size =12), 
        axis.title.x = element_text(family = "mono"),
        legend.text = element_text(family = "mono"),
        legend.title = element_text(family = "mono"))+
  scale_x_continuous(breaks = c(2010, 2012, 2014, 2016, 2018, 2020, 2022))+
  scale_color_natparks_d(name = "SmokyMtns")
```

