---
title: "Crime Data Cleaning"
author: "Erin Franke"
date: "2022-09-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the 4th file in my data cleaning and exploration process. \
Steps in this file: Download and clean Minneapolis and Saint Paul crime data.  \

## Data and library loading

```{r}
# Libraries
library(tidyverse)
library(sf)
library(stringr)

# Minneapolis crime 
minneapolisCrime2010 <- read_csv("Data/Crime/Police_Incidents_2010.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2011 <- read_csv("Data/Crime/Police_Incidents_2011.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2012 <- read_csv("Data/Crime/Police_Incidents_2012.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2013 <- read_csv("Data/Crime/Police_Incidents_2013.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2014 <- read_csv("Data/Crime/Police_Incidents_2014.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2015 <- read_csv("Data/Crime/Police_Incidents_2015.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2016 <- read_csv("Data/Crime/Police_Incidents_2016.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2017 <- read_csv("Data/Crime/Police_Incidents_2017.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2018a <- read_csv("Data/Crime/Police_Incidents_2018_PIMS.csv") %>%
  select(reportedDate, offense, description, centerLat, centerLong) %>%
  rename("ReportedDate" = "reportedDate", 
         "Offense" = "offense", 
         "Description" = "description", 
         "Lat" = "centerLat", 
         "Long" = "centerLong")
minneapolisCrime2018b <- read_csv("Data/Crime/Police_Incidents_2018.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2019 <- read_csv("Data/Crime/Police_Incidents_2019.csv") %>%
  select(reportedDate, offense, description, centerLat, centerLong) %>%
  rename("ReportedDate" = "reportedDate", 
         "Offense" = "offense", 
         "Description" = "description", 
         "Lat" = "centerLat", 
         "Long" = "centerLong")
minneapolisCrime2020 <- read_csv("Data/Crime/Police_Incidents_2020.csv") %>%
  select(reportedDate, offense, description, centerLat, centerLong) %>%
  rename("ReportedDate" = "reportedDate", 
         "Offense" = "offense", 
         "Description" = "description", 
         "Lat" = "centerLat", 
         "Long" = "centerLong")
minneapolisCrime2021 <- read_csv("Data/Crime/Police_Incidents_2021.csv") %>%
  select(reportedDate, offense, description, centerLat, centerLong) %>%
  rename("ReportedDate" = "reportedDate", 
         "Offense" = "offense", 
         "Description" = "description", 
         "Lat" = "centerLat", 
         "Long" = "centerLong")
minneapolisCrime2022 <- read_csv("Data/Crime/Police_Incidents_2022.csv") %>%
  select(reportedDate, offense, description, centerLat, centerLong) %>%
  rename("ReportedDate" = "reportedDate", 
         "Offense" = "offense", 
         "Description" = "description", 
         "Lat" = "centerLat", 
         "Long" = "centerLong")

# Saint Paul crime 
saintpaulcrime <- read_csv("Data/Crime/Crime_Incident_Report.csv")

# Gentrification Information
load('Data/gent_clean.RData')
```

## Minneapolis crime data

First, combine and organize crime files. 

```{r}
minneapolisCrime <- bind_rows(minneapolisCrime2010, minneapolisCrime2011, minneapolisCrime2012, minneapolisCrime2013, minneapolisCrime2014, minneapolisCrime2015, minneapolisCrime2016, minneapolisCrime2017, minneapolisCrime2018a, minneapolisCrime2018b, minneapolisCrime2019, minneapolisCrime2020, minneapolisCrime2021, minneapolisCrime2022) %>%
  mutate(year = as.numeric(substr(ReportedDate, 1, 4)), 
         date = as.Date.character(ReportedDate, format = "%Y/%m/%d"))
```

```{r, eval=FALSE}
save(minneapolisCrime, file = "Data/Crime/minneapolisCrime.RData")
```

```{r}
load("Data/Crime/minneapolisCrime.RData")
```

I want to aggregate these points by census tract based on latitude and longitude and plot them using the final census tracts from the cleaned gentrified file. 

Make the crime coordinates sf geometry points of the right coordinate system.
```{r}
minn_sf <- minneapolisCrime %>% 
  filter(Lat >0, Long <0) %>%
  st_as_sf(coords = c("Long", "Lat"), crs = "NAD83")
```

Aggregate the points by census tract. 
```{r}
tract_points <- st_join(minn_sf, left = FALSE, gent_clean["tract"]) 
plot(tract_points$geometry, pch=21, cex=0.7, col="purple", bg="gray80")
plot(gent_clean$geometry, border="gray20", col=NA, add=T)

crimes_per_tract <- tract_points %>%
  group_by(tract, year) %>%
  count()

tract_aggregated <- st_join(gent_clean, crimes_per_tract, left = FALSE) %>%
  select(-tract.y) %>%
  rename("tract" = "tract.x") %>%
  mutate(crimes_per_1000 = (n/population_2010) * 1000, 
         crimes_per_1000d = as.factor(case_when(crimes_per_1000 <=23 ~ "0-23", 
                                      crimes_per_1000 >23 & crimes_per_1000 <= 36 ~ "24-36", 
                                      crimes_per_1000 > 36 & crimes_per_1000 <= 52 ~ "37-52", 
                                      crimes_per_1000 >52 & crimes_per_1000 <= 75 ~ "53-75", 
                                      TRUE ~ "75+")), 
         crimes_per_1000d = fct_relevel(crimes_per_1000d, levels = c("0-23", "24-36", "37-52", "53-75", "75+"))) %>%
  filter(!(tract == 1055 & year == 2012 & n ==150), 
         !(tract == 1066 & year == 2012 & n ==181))
```

Join crime data with `gent_clean` file.

```{r}
tract_points <- st_join(minn_sf, left = FALSE, gent_clean["tract"]) %>%
  st_join(gent_clean) %>%
  select(-tract.y) %>%
  rename("tract" = "tract.x")
```


```{r,eval=FALSE}
save(tract_aggregated, file = 'Data/tract_aggregated.RData')
save(tract_points, file = "Data/tract_points.RData")
```

Specifically look at violent crimes

```{r}
crimes_per_tract_violent <- tract_points %>%
  mutate(offense_group = case_when(Offense %in% c("THEFT", "BIKETF", "TFPER", "THFTSW", "PKGTHEFT") ~ "Theft", 
                                   Offense %in% c("TFMV", "AUTOTH", "TMVP", "CARJCK", "MVTHFT") ~ "Auto Theft",
                                   Offense == "BURGD" ~ "Home Burglary", 
                                   Offense %in% c("BURGB", "ROBBIZ", "TBLDG") ~ "Business Burglary", 
                                   Offense %in% c("ASLT2", "CSCR", "DASTR", "DASLT2", "ASLT3", "ASLT4", "DASLT3", "MURDR", "ASLT1", "DASLT1") ~ "Assault", 
                                   TRUE ~ "Other")) %>%
  filter(offense_group == "Assault") %>%
  group_by(tract, year) %>%
  count()

tract_aggregated_violent <- st_join(gent_clean, crimes_per_tract_violent, left = FALSE) %>%
  select(-tract.y) %>%
  rename("tract" = "tract.x") %>%
  mutate(crimes_per_1000 = (n/population_2010) * 1000)
```

Specifically look at thefts

```{r}
crimes_per_tract_theft <- tract_points %>%
  mutate(offense_group = case_when(Offense %in% c("THEFT", "BIKETF", "TFPER", "THFTSW", "PKGTHEFT") ~ "Theft", 
                                   Offense %in% c("TFMV", "AUTOTH", "TMVP", "CARJCK", "MVTHFT") ~ "Auto Theft",
                                   Offense == "BURGD" ~ "Home Burglary", 
                                   Offense %in% c("BURGB", "ROBBIZ", "TBLDG") ~ "Business Burglary", 
                                   Offense %in% c("ASLT2", "CSCR", "DASTR", "DASLT2", "ASLT3", "ASLT4", "DASLT3", "MURDR", "ASLT1", "DASLT1") ~ "Assault", 
                                   TRUE ~ "Other")) %>%
  filter(offense_group %in% c("Theft", "Auto Theft")) %>%
  group_by(tract, year) %>%
  count()

tract_aggregated_theft <- st_join(gent_clean, crimes_per_tract_theft, left = FALSE) %>%
  select(-tract.y) %>%
  rename("tract" = "tract.x") %>%
  mutate(crimes_per_1000 = (n/population_2010) * 1000) %>%
  filter(!(tract == 1055 & year == 2012 & n ==104), 
         !(tract == 1066 & year == 2012 & n ==110))
```

```{r, eval=FALSE}
save(tract_aggregated_violent, file = "Data/Crime/tract_aggregated_violent.RData")
save(tract_aggregated_theft, file = "Data/Crime/tract_aggregated_theft.RData")
```


## Cleaning saint paul crime data

Unlike the Minneapolis crime data, the Saint Paul data did not come with a latitude and longitude location of each crime. However, it did come with a rough block address or street intersection. Using GIS Pro we were able to get a rough latitude and longitude of each crime and then follow the same process as the Minneapolis data.

### Preparing the data for GIS 

First, identify crimes that occurred at intersections. These will need to be treated differently than those with addresses. Use string techniques to put them in a format that GIS likes. 
```{r}
intersections <- saintpaulcrime %>%
  select(BLOCK) %>%
  filter(str_detect(BLOCK, "&")) %>%
  count(BLOCK) %>%
  select(-n)

streets <- as.data.frame(str_split_fixed(intersections$BLOCK, ' & ', 2))
street_identifiers <- as.data.frame(str_split_fixed(streets$V1, " ", 2)) %>%
  #mutate(V2 = word(V2, 1)) %>% trying this out
  rename("street" = "V1", 
         "identifier" = "V2")
  
intersections_RamseyCo <- as.data.frame(str_split_fixed(intersections$BLOCK, ' & ', 2)) %>%
  left_join(street_identifiers, by = c("V2" = "street")) %>%
  mutate(V2 = paste(V2, identifier, sep= " ")) %>%
  select(-identifier) %>%
  rename("street1" = "V1", 
         "street2" = "V2") %>%
  mutate(street1 = str_trim(street1, side = c("right")), 
         street2 = str_trim(street2, side = "right")) %>%
  distinct_all() %>%
  mutate(combined = paste(street1, street2, sep = " & ")) %>%
  select(combined)
```

```{r, eval=FALSE}
write_csv(intersections_RamseyCo, "Data/Crime/intersections_RamseyCo.csv")
```

### Cleaning up GIS data 

Import data and select rough intersection and coordinates.
```{r}
intersections_ll <- read_csv("Data/Crime/intersections_RamseyCo_geocoded2.csv")
matched_lat_lngs <- intersections_ll %>%
  filter(Score > 0) %>%
  select(USER_combined, X, Y)
```

Create same intersection descriptions and make column `original` so we are able to join back to original data and get crime date and type. 
```{r}
all_intersections <- saintpaulcrime %>%
  filter(str_detect(BLOCK, "&")) %>%
  select(DATE, INCIDENT, BLOCK)

streets <- as.data.frame(str_split_fixed(all_intersections$BLOCK, ' & ', 2)) 
streets2 <- bind_cols(streets, all_intersections)
street_identifiers <- as.data.frame(str_split_fixed(streets2$V1, " ", 2)) %>%
  mutate(V2 = word(V2, 1)) %>% 
  rename("street" = "V1", 
         "identifier" = "V2")
  
intersections_clean <- streets2 %>%
  left_join(street_identifiers, by = c("V2" = "street")) %>%
  distinct() %>%
  mutate(V2 = paste(V2, identifier, sep= " ")) %>%
  select(-identifier) %>%
  rename("street1" = "V1", 
         "street2" = "V2") %>%
  mutate(street1 = str_trim(street1, side = c("right")), 
         street2 = str_trim(street2, side = "right")) %>%
  mutate(combined = paste(street1, street2, sep = " & ")) 
```

```{r}
intersections_clean %>%
  left_join(matched_lat_lngs, by = c("combined" = "USER_combined")) %>%
  distinct(DATE, X, Y, INCIDENT) %>%
  na.omit()
```

### Estimating addresses when given the block

```{r}
blocks <- saintpaulcrime %>%
  filter(!str_detect(BLOCK, "&")) %>%
  select(DATE, INCIDENT, BLOCK) %>%
  mutate(BLOCK = str_replace(BLOCK, "X", "0")) %>%
  tidygeocoder::geocode(BLOCK, method = 'osm', lat = latitude, long = longitude)

do_one_sim_three <- function(i){
  # simulate null trait
  blocks <- saintpaulcrime %>%
  filter(!str_detect(BLOCK, "&")) %>%
  select(DATE, INCIDENT, BLOCK) %>%
  mutate(BLOCK = str_replace(BLOCK, "X", "0")) %>%
  distinct(BLOCK)
  address_lat_long = data.frame(address = c(), lat = c(), long = c())
  # implement GWAS
  for(i in 1:32562){
    address_lat_long$address[i] <- blocks$BLOCK
    address_lat_long$latitude[i] <- tidygeocoder::geocode(blocks$BLOCK %>% as.data.frame(), method = 'osm', lat = latitude, long = longitude)$latitude
    address_lat_lpng$longitude[i] <- tidygeocoder::geocode(blocks$BLOCK %>% as.data.frame(), method = 'osm', lat = latitude, long = longitude)$longitude
  }
}

mclapply(1, do_one_sim_three, mc.cores = 4)
```

