---
title: "Crime Data Cleaning"
author: "Erin Franke"
date: "2022-09-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the 5th file in my data cleaning and exploration process. \
Steps in this file: Download and clean Minneapolis and Saint Paul crime data.  \

## Data and library loading

```{r}
# Libraries
library(tidyverse)
library(sf)
library(stringr)
library(NatParksPalettes)
```

Create Minneapolis crime file from 2010 to September 2022.

```{r, eval=FALSE}
# Minneapolis crime 
minneapolisCrime2010 <- read_csv("Data/Crime/Police_Incidents_2010.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2011 <- read_csv("Data/Crime/Police_Incidents_2011.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2012 <- read_csv("Data/Crime/Police_Incidents_2012.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2013 <- read_csv("Data/Crime/Police_Incidents_2013.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2014 <- read_csv("Data/Crime/Police_Incidents_2014.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2015 <- read_csv("Data/Crime/Police_Incidents_2015.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2016 <- read_csv("Data/Crime/Police_Incidents_2016.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2017 <- read_csv("Data/Crime/Police_Incidents_2017.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2018a <- read_csv("Data/Crime/Police_Incidents_2018_PIMS.csv") %>%
  select(reportedDate, offense, description, centerLat, centerLong) %>%
  rename("ReportedDate" = "reportedDate", 
         "Offense" = "offense", 
         "Description" = "description", 
         "Lat" = "centerLat", 
         "Long" = "centerLong")
minneapolisCrime2018b <- read_csv("Data/Crime/Police_Incidents_2018.csv") %>%
  select(ReportedDate, Offense, Description, Lat, Long)
minneapolisCrime2019 <- read_csv("Data/Crime/Police_Incidents_2019.csv") %>%
  select(reportedDate, offense, description, centerLat, centerLong) %>%
  rename("ReportedDate" = "reportedDate", 
         "Offense" = "offense", 
         "Description" = "description", 
         "Lat" = "centerLat", 
         "Long" = "centerLong")
minneapolisCrime2020 <- read_csv("Data/Crime/Police_Incidents_2020.csv") %>%
  select(reportedDate, offense, description, centerLat, centerLong) %>%
  rename("ReportedDate" = "reportedDate", 
         "Offense" = "offense", 
         "Description" = "description", 
         "Lat" = "centerLat", 
         "Long" = "centerLong")
minneapolisCrime2021 <- read_csv("Data/Crime/Police_Incidents_2021.csv") %>%
  select(reportedDate, offense, description, centerLat, centerLong) %>%
  rename("ReportedDate" = "reportedDate", 
         "Offense" = "offense", 
         "Description" = "description", 
         "Lat" = "centerLat", 
         "Long" = "centerLong")
minneapolisCrime2022 <- read_csv("Data/Crime/Police_Incidents_2022.csv") %>%
  select(reportedDate, offense, description, centerLat, centerLong) %>%
  rename("ReportedDate" = "reportedDate", 
         "Offense" = "offense", 
         "Description" = "description", 
         "Lat" = "centerLat", 
         "Long" = "centerLong")

minneapolisCrime <- bind_rows(minneapolisCrime2010, minneapolisCrime2011, minneapolisCrime2012, minneapolisCrime2013, minneapolisCrime2014, minneapolisCrime2015, minneapolisCrime2016, minneapolisCrime2017, minneapolisCrime2018a, minneapolisCrime2018b, minneapolisCrime2019, minneapolisCrime2020, minneapolisCrime2021, minneapolisCrime2022) %>%
  mutate(year = as.numeric(substr(ReportedDate, 1, 4)), 
         date = as.Date.character(ReportedDate, format = "%Y/%m/%d"))

save(minneapolisCrime, file = "Data/Crime/minneapolisCrime.RData")
```

Load Saint Paul crime, Minneapolis crime, and gentrification file.

```{r}
# Saint Paul crime 
saintpaulcrime <- read_csv("Data/Crime/Crime_Incident_Report.csv")

# Minneapolis Crime
load("Data/Crime/minneapolisCrime.RData")

# Gentrification Information
load('Data/tractData/cleaned/gent_clean.RData')
```

## Clean Minneapolis crime data

Goal: aggregate these points by census tract based on latitude and longitude and plot them using the final census tracts from the cleaned gentrification file. 

Make the crime coordinates sf geometry points of the right coordinate system.

```{r}
minn_sf <- minneapolisCrime %>% 
  filter(Lat >0, Long <0) %>%
  st_as_sf(coords = c("Long", "Lat"), crs = "NAD83") %>%
  distinct()
```

Aggregate the points by census tract. 

```{r}
tract_points <- st_join(minn_sf, left = FALSE, gent_clean["tractmod"]) 
plot(tract_points$geometry, pch=21, cex=0.7, col="purple", bg="gray80")
plot(gent_clean$geometry, border="gray20", col=NA, add=T)

crimes_per_tract <- tract_points %>%
  group_by(tractmod, year) %>%
  count()

tract_aggregated <- st_join(gent_clean, crimes_per_tract, left = FALSE) %>%
  select(-tractmod.y) %>%
  rename("tract" = "tractmod.x") %>%
  mutate(crimes_per_1000 = (n/population_2020) * 1000, 
         crimes_per_1000d = as.factor(case_when(crimes_per_1000 <=23 ~ "0-23", 
                                      crimes_per_1000 >23 & crimes_per_1000 <= 36 ~ "24-36", 
                                      crimes_per_1000 > 36 & crimes_per_1000 <= 52 ~ "37-52", 
                                      crimes_per_1000 >52 & crimes_per_1000 <= 75 ~ "53-75", 
                                      TRUE ~ "75+")), 
         crimes_per_1000d = fct_relevel(crimes_per_1000d, levels = c("0-23", "24-36", "37-52", "53-75", "75+"))) %>%
  filter(!(tract == 1055 & year == 2012 & n ==150), 
         !(tract == 1066 & year == 2012 & n ==181))
```

Join crime data with `gent_clean` file.

```{r}
tract_points <- st_join(minn_sf, left = FALSE, gent_clean["tractmod"]) %>%
  st_join(gent_clean) %>%
  select(-tractmod.y) %>%
  rename("tract" = "tractmod.x") %>%
  mutate(offense_group = case_when(Offense %in% c("THEFT", "BIKETF", "TFPER", "THFTSW", "PKGTHEFT") ~ "Theft", 
                                   Offense %in% c("TFMV", "AUTOTH", "TMVP", "CARJCK", "MVTHFT") ~ "Auto Theft",
                                   Offense == "BURGD" ~ "Home Burglary", 
                                   Offense %in% c("BURGB", "ROBBIZ", "TBLDG") ~ "Business Burglary", 
                                   Offense %in% c("ASLT2", "CSCR", "DASTR", "DASLT2", "ASLT3", "ASLT4", "DASLT3", "MURDR", "ASLT1", "DASLT1") ~ "Assault", 
                                   TRUE ~ "Other"))
```

```{r,eval=FALSE}
save(tract_aggregated, file = 'Data/CrimeCleaned/tract_aggregated_minneapolis.RData')
save(tract_points, file = "Data/CrimeCleaned/tract_points_minneapolis.RData")
```

### Violent crimes in Minneapolis

```{r}
crimes_per_tract_violent <- tract_points %>%
  filter(offense_group == "Assault") %>%
  group_by(tract, year) %>%
  count()

tract_aggregated_violent <- st_join(gent_clean, crimes_per_tract_violent %>% rename("tractmod" = "tract"), left = FALSE) %>%
  select(-tractmod.y) %>%
  rename("tractmod" = "tractmod.x") %>%
  mutate(crimes_per_1000 = (n/population_2020) * 1000)
```

### Thefts in Minneapolis

```{r}
crimes_per_tract_theft <- tract_points %>%
  filter(offense_group == "Theft") %>%
  group_by(tract, year) %>%
  count()

tract_aggregated_theft <- st_join(gent_clean, crimes_per_tract_theft %>% rename("tractmod" = "tract"), left = FALSE) %>%
  select(-tractmod.y) %>%
  rename("tractmod" = "tractmod.x") %>%
  mutate(crimes_per_1000 = (n/population_2020) * 1000)
```

### Auto Thefts in Minneapolis

```{r}
crimes_per_tract_autotheft <- tract_points %>%
  filter(offense_group == "Auto Theft") %>%
  group_by(tract, year) %>%
  count()

tract_aggregated_autotheft <- st_join(gent_clean, crimes_per_tract_theft %>% rename("tractmod" = "tract"), left = FALSE) %>%
  select(-tractmod.y) %>%
  rename("tractmod" = "tractmod.x") %>%
  mutate(crimes_per_1000 = (n/population_2020) * 1000)
```


```{r, eval=FALSE}
save(tract_aggregated_violent, file = "Data/CrimeCleaned/tract_aggregated_minneapolis_violent.RData")
save(tract_aggregated_theft, file = "Data/CrimeCleaned/tract_aggregated_minneapolis_theft.RData")
save(tract_aggregated_autotheft, file = "Data/CrimeCleaned/tract_aggregated_minneapolis_autotheft.RData")
```

## Cleaning saint paul crime data

Unlike the Minneapolis crime data, the Saint Paul data did not come with a latitude and longitude location of each crime. However, it did come with a rough block address or street intersection. Using GIS Pro we were able to get a rough latitude and longitude of each crime and then follow the same process as the Minneapolis data.

### Preparing the data for GIS 

First, identify crimes that occurred at intersections. These will need to be treated differently than those with addresses. Use string techniques to put them in a format that GIS likes. 

```{r, eval=FALSE}
intersections <- saintpaulcrime %>%
  select(BLOCK) %>%
  filter(str_detect(BLOCK, "&")) %>%
  count(BLOCK) %>%
  select(-n)

streets <- as.data.frame(str_split_fixed(intersections$BLOCK, ' & ', 2))
street_identifiers <- as.data.frame(str_split_fixed(streets$V1, " ", 2)) %>%
  #mutate(V2 = word(V2, 1)) %>% trying this out
  rename("street" = "V1", 
         "identifier" = "V2")
  
intersections_RamseyCo <- as.data.frame(str_split_fixed(intersections$BLOCK, ' & ', 2)) %>%
  left_join(street_identifiers, by = c("V2" = "street")) %>%
  mutate(V2 = paste(V2, identifier, sep= " ")) %>%
  select(-identifier) %>%
  rename("street1" = "V1", 
         "street2" = "V2") %>%
  mutate(street1 = str_trim(street1, side = c("right")), 
         street2 = str_trim(street2, side = "right")) %>%
  distinct_all() %>%
  mutate(combined = paste(street1, street2, sep = " & ")) %>%
  select(combined)
```

```{r, eval=FALSE}
write_csv(intersections_RamseyCo, "Data/Crime/intersections_RamseyCo.csv")
```

### Cleaning up GIS data 

Import data and select rough intersection and coordinates.

```{r}
intersections_ll <- read_csv("Data/Crime/intersections_RamseyCo_geocoded2.csv")
matched_lat_lngs <- intersections_ll %>%
  filter(Score > 0) %>%
  select(USER_combined, X, Y)
```

Create same intersection descriptions and make column `original` so we are able to join back to original data and get crime date and type. 

```{r}
all_intersections <- saintpaulcrime %>%
  filter(str_detect(BLOCK, "&")) %>%
  select(DATE, INCIDENT, BLOCK)

streets <- as.data.frame(str_split_fixed(all_intersections$BLOCK, ' & ', 2)) 
streets2 <- bind_cols(streets, all_intersections)
street_identifiers <- as.data.frame(str_split_fixed(streets2$V1, " ", 2)) %>%
  mutate(V2 = word(V2, 1)) %>% 
  rename("street" = "V1", 
         "identifier" = "V2")

#problem: streets have different identifiers, causing multiple crimes
street_identifiers %>%
  distinct() %>%
  group_by(street) %>%
  mutate(count = n()) %>%
  filter(count>=2) %>%
  arrange(desc(count), street)

#best solution: take the most often identifier and remove other identifiers. This will make certain areas slightly short of crimes but prevents extreme double/triple/quadrouple counting.
street_identifiers2 <- street_identifiers %>%
  group_by(street) %>%
  count(identifier) %>%
  mutate(most_often_identifier = max(n)) %>%
  filter(most_often_identifier == n) %>%
  select(-n, -most_often_identifier)

intersections_clean <- streets2 %>%
  left_join(street_identifiers2, by = c("V2" = "street")) %>%
  distinct() %>%
  mutate(V2 = paste(V2, identifier, sep= " ")) %>%
  select(-identifier) %>%
  rename("street1" = "V1", 
         "street2" = "V2") %>%
  mutate(street1 = str_trim(street1, side = c("right")), 
         street2 = str_trim(street2, side = "right")) %>%
  mutate(combined = paste(street1, street2, sep = " & ")) 
```

```{r}
intersection_crimes <- intersections_clean %>%
  left_join(matched_lat_lngs, by = c("combined" = "USER_combined")) %>%
  distinct(DATE, X, Y, INCIDENT) %>%
  na.omit() %>%
  rename("longitude" = "X", 
         "latitude" = "Y") %>%
  mutate(year = as.numeric(substr(DATE,1, 4)))
```

### Estimating addresses when given the block

I used tidygeocoder to get estimated coordinates given the block the crime occurred. The following code chunk took close to 10 hours to run. 

```{r, eval=FALSE}
blocks <- saintpaulcrime %>%
  filter(!str_detect(BLOCK, "&")) %>%
  select(DATE, INCIDENT, BLOCK) %>%
  mutate(BLOCK = str_replace(BLOCK, "X", "0")) %>%
  tidygeocoder::geocode(BLOCK, method = 'osm', lat = latitude, long = longitude)
save(blocks, file = "Data/Crime/address_latlngs.RData")
```

Load the addresses from above. Some addresses were not able to get coordinates - a large reason for this is because the data combined streets with two names (for example St Clair was STCLAIR). 

```{r}
load("Data/Crime/address_latlngs.RData")
  
#address with recorded latitude and longitude (87.7%)
address_clean <- blocks %>%
  select(-BLOCK) %>%
  mutate(year = as.numeric(substr(DATE,1, 4))) %>%
  na.omit()
```

Filter for addresses that did not get an identified latitude and longitude. There are 36,149 and about half can hopefully be fixed by adding a space to the two named streets.

```{r}
two_named_streets <- blocks %>%
  filter(is.na(latitude)) %>%
  separate(col = BLOCK, into = c("number", "street", "identifier", "direction"), sep = " ") %>%
  mutate(street2 = case_when(str_detect(street, "OLDHUDSON") ~ "OLD HUDSON",
                            str_detect(street, "ANNARBOR") ~ "ANN ARBOR",
                            str_detect(street, "BARGECHANNEL") ~ "BARGE CHANNEL",
                            str_detect(street, "BATTLECREEK") ~ "BATTLE CREEK",
                            str_detect(street, "CESARCHAVEZ") ~ "CESAR CHAVEZ",
                            str_detect(street, "CHEROKEEHTS") ~ "CHEROKEE HTS", 
                            str_detect(street, "CROCUSHILL") ~ "CROCUS HILL", 
                            str_detect(street, "CROSBYFARM") ~ "CROSBY FARM",
                            str_detect(street, "CROSBYLAKE") ~ "CROSBY LAKE",
                            str_detect(street, "DANPATCH") ~ "DAN PATCH", 
                            str_detect(street, "DOROTHYDAY") ~ "DOROTHY DAY",
                            str_detect(street, "DRJUSTUSOHAGE") ~ "DR JUSTUS OHAGE", 
                            str_detect(street, "EDGCUMBE") ~ "EDGECUMBE", 
                            str_detect(street, "ENERGYPARK") ~ "ENERGY PARK",
                            str_detect(street, "GRANDHILL") ~ "GRAND HILL",
                            str_detect(street, "HARDENBERGH") ~ "HARDEN BERGH", 
                            str_detect(street, "HARRIETISLAND") ~ "HARRIET ISLAND", 
                            str_detect(street, "HIDDENFALLS") ~ "HIDDEN FALLS",
                            str_detect(street, "HUNTINGVALLEY") ~ "HUNTING VALLEY", 
                            str_detect(street, "IVANWAY") ~ "IVAN WAY", 
                            str_detect(street, "JEFFERSONCOMMONS") ~ "JEFFERSON COMMONS",
                            str_detect(street, "JOHNIRELAND") ~ "JOHN IRELAND", 
                            str_detect(street, "LAFAYETTEFRONTAGE") ~ "LAFAYETTE FRONTAGE",
                            str_detect(street, "LARRYHO") ~ "LARRY HO", 
                            str_detect(street, "LOWERAFTON") ~ "LOWER AFTON", 
                            str_detect(street, "LOWERSTDENNIS") ~ "LOWER ST DENNIS", 
                            str_detect(street, "LYNNHURST") ~ "LYNN HURST", 
                            str_detect(street, "MARTINLUTHERKING") ~ "MARTIN LUTHER KING", 
                            str_detect(street, "MISSISSIPPIRIVER") ~ "MISSISSIPPI RIVER", 
                            str_detect(street, "MORNINGSIDE") ~ "MORNING SIDE",
                            str_detect(street, "MTAIRY") ~ "MT AIRY", 
                            str_detect(street, "MTCURVE") ~ "MT CURVE",
                            str_detect(street, "OLDSIXTH") ~ "OLD SIXTH",
                            str_detect(street, "OTTOHUMMER") ~ "OTTO HUMMER",
                            str_detect(street, "PIERCEBUTLER") ~ "PIERCE BUTLER",
                            str_detect(street, "PIGSEYELAKE") ~ "PIGSEYE LAKE",
                            str_detect(street, "PTDOUGLAS") ~ "PT DOUGLAS",
                            str_detect(street, "RIVERPARK") ~ "RIVER PARK",
                            str_detect(street, "SPRUCETREE") ~ "SPRUCE TREE",
                            str_detect(street, "STANTHONY") ~ "ST ANTHONY",
                            str_detect(street, "STALBANS") ~ "ST ALBANS",
                            str_detect(street, "STCLAIR") ~ "ST CLAIR",
                            str_detect(street, "STDENNIS") ~ "ST DENNIS",
                            str_detect(street, "STPAUL") ~ "ST PAUL",
                            str_detect(street, "STPETER") ~ "ST PETER",
                            str_detect(street, "TROUTBROOK") ~ "TROUT BROOK",
                            str_detect(street, "UPPERAFTON") ~ "UPPER AFTON",
                            str_detect(street, "UPPERSTDENNIS") ~ "UPPER ST DENNIS",
                            str_detect(street, "VANBUREN") ~ "VAN BUREN",
                            str_detect(street, "VANSLYKE") ~ "VAN SLYKE",
                            str_detect(street, "WHEELOCKRIDGE") ~ "WHEELOCK RIDGE",
                            str_detect(street, "WHITEBEAR") ~ "WHITE BEAR",
                            str_detect(street, "WILLIAMTELL") ~ "WILLIAM TELL",
                            str_detect(street, "YACHTCLUB") ~ "YACHT CLUB", 
                            TRUE ~ street)) %>%
  filter(street2 != street) %>%
  mutate(identifier = case_when(is.na(identifier) ~ "", 
                                TRUE ~ identifier), 
         direction = case_when(is.na(direction) ~ "", 
                               TRUE ~ direction)) %>%
  mutate(new_address = paste(number, street2, identifier, direction, sep = " "), 
         new_address = str_trim(new_address, side = c("right")))
```

Run the two named streets through the geocoder. This gets us almost 16,000 of the 36,000+ missing crimes.

```{r, eval=FALSE}
blocks_two_name <- two_named_streets %>%
  tidygeocoder::geocode(new_address, method = 'osm', lat = latitude, long = longitude)

blocks_two_name <- blocks_two_name %>%
  filter(!is.na(`latitude...11`)) %>%
  select(-number, -street, -identifier, -direction, -`latitude...7`, -`longitude...8`, -street2) %>%
  rename("latitude" = "latitude...11", 
         "longitude" = "longitude...12")

save(blocks_two_name, file = "Data/CrimeCleaned/blocks_two_name.RData")
```

For street names with a large amount of data points that are not two worded streets, identify them and try to figure out what is wrong. Not all of these can be fixed, but I tried to find a few.

```{r}
# identify more common problems (some of these are not easily fixable or already fixed)
 blocks %>%
  filter(is.na(latitude)) %>%
  separate(col = BLOCK, into = c("number", "street", "identifier", "direction"), sep = " ") %>%
  count(street, identifier) %>%
  arrange(desc(n)) %>%
  head(60)
```

From the list above, fix a few streets that were having issues due to the identifier abbreviation. 

```{r, eval=FALSE}
# try replacing bd with blvd 
blocks_blvd <- blocks %>%
  filter(is.na(latitude)) %>%
  separate(col = BLOCK, into = c("number", "street", "identifier", "direction"), sep = " ", remove = FALSE) %>%
  mutate(street = case_when(street == "CHEROKEEHTS" ~ "CHEROKEE HTS",
                            street == "MTCURVE" ~ "MT CURVE",
                            TRUE ~ street)) %>%
   filter(street %in% c("CAPITOL", "BANDANA", "KELLOGG", "PHALEN", "CHEROKEE HTS", "PLATO", "MT CURVE", "COMO")) %>%
   mutate(identifier = "BLVD") %>%
   mutate(identifier = case_when(is.na(identifier) ~ "", 
                                TRUE ~ identifier), 
          direction = case_when(is.na(direction) ~ "", 
                               TRUE ~ direction),
          new_address = paste(number, street, identifier, direction, sep = " "), 
          new_address = str_trim(new_address, side = c("right"))) %>%
   select(-latitude, -longitude) %>%
   tidygeocoder::geocode(new_address, method = 'osm', lat = latitude, long = longitude) %>%
  na.omit(latitude)

blocks_pkwy <- blocks %>%
  filter(is.na(latitude)) %>%
  separate(col = BLOCK, into = c("number", "street", "identifier", "direction"), sep = " ", remove = FALSE) %>%
   filter(street %in% c("LEXINGTON", "WHEELOCK")) %>%
   mutate(identifier = "PKWY") %>%
   mutate(identifier = case_when(is.na(identifier) ~ "", 
                                TRUE ~ identifier), 
          direction = case_when(is.na(direction) ~ "", 
                               TRUE ~ direction),
          new_address = paste(number, street, identifier, direction, sep = " "), 
          new_address = str_trim(new_address, side = c("right"))) %>%
   select(-latitude, -longitude) %>%
   tidygeocoder::geocode(new_address, method = 'osm', lat = latitude, long = longitude) %>%
  na.omit(latitude)

blocks_blvd <- blocks_blvd %>%
  select(DATE, INCIDENT, new_address, latitude, longitude)
blocks_pkwy <- blocks_pkwy %>%
  select(DATE, INCIDENT, new_address, latitude, longitude)

save(blocks_blvd, file = "Data/CrimeCleaned/blocks_blvd.RData")
save(blocks_pkwy, file = "Data/CrimeCleaned/blocks_pkwy.RData")
```

Finally, fix the numbered streets.

```{r, eval=FALSE}
numbered_streets <- saintpaulcrime %>%
  filter(!str_detect(BLOCK, "&")) %>%
  select(DATE, INCIDENT, BLOCK) %>%
  mutate(containsXX = case_when(str_detect(BLOCK, "XX") ~ 1, 
         TRUE ~ 0)) %>%
  filter(containsXX == 1) %>%
  separate(col = BLOCK, into = c("number", "street", "identifier", "direction"), sep = " ", remove = FALSE) %>%
  filter(str_length(number) == 2) %>%
  mutate(number = sample(c(10, 20, 30, 40, 50, 60, 70, 80, 90), 2026, replace=TRUE)) %>%
  mutate(identifier = case_when(is.na(identifier) ~ "", 
                                TRUE ~ identifier), 
          direction = case_when(is.na(direction) ~ "", 
                               TRUE ~ direction),
          new_address = paste(number, street, identifier, direction, sep = " "), 
          new_address = str_trim(new_address, side = c("right"))) %>%
  select(-BLOCK, -number, -street, -identifier, -direction, -containsXX) %>%
  tidygeocoder::geocode(new_address, method = 'osm', lat = latitude, long = longitude)

numbered_streets <- numbered_streets %>%
  filter(!is.na(latitude))

save(numbered_streets, file = "Data/CrimeCleaned/numbered_streets.RData")
```


### Combine intersection and block crimes 

Create violent, theft, and autotheft data.
Clean up data repository
Crime EDA 
Crime modeling

Combine cleaned files to get Saint Paul crimes.

```{r, eval=FALSE}
load("Data/CrimeCleaned/numbered_streets.RData")
load("Data/CrimeCleaned/blocks_blvd.RData")
load("Data/CrimeCleaned/blocks_pkwy.RData")

numbered_streets <- numbered_streets %>%
  mutate(year = as.numeric(substr(DATE, 1, 4))) %>%
  select(DATE, INCIDENT, latitude, longitude, year)
blocks_blvd <- blocks_blvd %>%
  mutate(year = as.numeric(substr(DATE, 1, 4))) %>%
  select(DATE, INCIDENT, latitude, longitude, year)
blocks_pkwy <- blocks_pkwy %>%
  mutate(year = as.numeric(substr(DATE, 1, 4))) %>%
  select(DATE, INCIDENT, latitude, longitude, year)
  

sp_crimes_located <- bind_rows(address_clean, intersection_crimes, numbered_streets, blocks_blvd, blocks_pkwy)
save(sp_crimes_located, file = "Data/CrimeCleaned/sp_crimes_located.RData")
```

Make the crime coordinates sf geometry points of the right coordinate system.

```{r}
sp_sf <- sp_crimes_located %>%
  filter(latitude >0, longitude<0) %>%
  distinct() %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = "NAD83")
```

Aggregate the points by census tract. 

```{r}
tract_points_sp <- st_join(sp_sf, left = FALSE, gent_clean["tractmod"]) 
plot(tract_points_sp$geometry, pch=21, cex=0.7, col="purple", bg="gray80")
plot(gent_clean$geometry, border="gray20", col=NA, add=T)

crimes_per_tract_sp <- tract_points_sp %>%
  filter(!(INCIDENT %in% c("Proactive Police Visit", "Proactive Foot Patrol", "Community Engagement Event", "Community Event", "Discharge")), !is.na(INCIDENT)) %>%
  group_by(tractmod, year) %>%
  count()

tract_aggregated_sp <- st_join(gent_clean, crimes_per_tract_sp, left = FALSE) %>%
  select(-tractmod.y) %>%
  rename("tractmod" = "tractmod.x") %>%
  mutate(crimes_per_1000 = (n/population_2020) * 1000, 
         crimes_per_1000d = as.factor(case_when(crimes_per_1000 <=23 ~ "0-23", 
                                      crimes_per_1000 >23 & crimes_per_1000 <= 36 ~ "24-36", 
                                      crimes_per_1000 > 36 & crimes_per_1000 <= 52 ~ "37-52", 
                                      crimes_per_1000 >52 & crimes_per_1000 <= 75 ~ "53-75", 
                                      TRUE ~ "75+")), 
         crimes_per_1000d = fct_relevel(crimes_per_1000d, levels = c("0-23", "24-36", "37-52", "53-75", "75+"))) %>%
  filter(county_2020 == "Ramsey")
```

Join crime data with `gent_clean` file.

```{r}
tract_points_sp <- st_join(sp_sf, left = FALSE, gent_clean["tractmod"]) %>%
  st_join(gent_clean) %>%
  select(-tractmod.y) %>%
  rename("tractmod" = "tractmod.x") %>%
  filter(!(INCIDENT %in% c("Proactive Police Visit", "Proactive Foot Patrol", "Community Engagement Event", "Community Event", "Discharge")), !is.na(INCIDENT)) %>%
  mutate(incident_group = case_when(INCIDENT %in% c("THEFT", "Theft") ~ "Theft", 
                                   INCIDENT %in% c("Auto Theft") ~ "Auto Theft",
                                   INCIDENT %in% c("Burglary", "Robbery") ~ "Burglary",
                                   INCIDENT %in% c("Agg. Assault", "Agg. Assault Dom", "Agg. Assault Dom.", "Homicide", "Rape", "Simple Asasult Dom.", "Simple Assault Dom", "Simple Assault Dom.") ~ "Assault", 
                                   TRUE ~ "Other"))
```


```{r,eval=FALSE}
save(tract_aggregated_sp, file = 'Data/CrimeCleaned/tract_aggregated_sp.RData')
save(tract_points_sp, file = "Data/CrimeCleaned/tract_points_sp.RData")
```

