---
title: "Crime Data Cleaning"
author: "Erin Franke"
date: "2022-09-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Minneapolis crime data

I want to aggregate these points by census tract based on latitude and longitude and plot them using the final census tracts from the cleaned gentrified file. 

```{r}
minneapolisCrime2022 <- read_csv("Data/Crime/Police_Incidents_2022.csv")
saintpaulcrime <- read_csv("Data/Crime/Crime_Incident_Report.csv")
load('Data/gent_clean.RData')
library(NatParksPalettes)
library(PrettyCols)
library(ggmap)
```

Make the crime coordinates sf geometry points of the right coordinate system.
```{r}
minn_sf <- minneapolisCrime2022 %>% 
  filter(X < 0) %>%
  st_as_sf(coords = c("X", "Y"), crs = "NAD83") %>%
  select(offense)
```

Aggregate the points by census tract. 
```{r}
tract_points <- st_join(minn_sf, left = FALSE, gent_clean["tract"]) 
plot(tract_points$geometry, pch=21, cex=0.7, col="purple", bg="gray80")
plot(gent_clean$geometry, border="gray20", col=NA, add=T)

crimes_per_tract <- tract_points %>%
  group_by(tract) %>%
  count()

tract_aggregated_2022 <- st_join(gent_clean, crimes_per_tract, left = FALSE) %>%
  select(-tract.y) %>%
  rename("tract" = "tract.x") %>%
  mutate(crimes_per_1000 = (n/population_2010) * 1000, 
         crimes_per_1000d = as.factor(case_when(crimes_per_1000 <=23 ~ "0-23", 
                                      crimes_per_1000 >23 & crimes_per_1000 <= 37 ~ "24-37", 
                                      crimes_per_1000 > 37 & crimes_per_1000 <= 50 ~ "38-50", 
                                      crimes_per_1000 >50 & crimes_per_1000 <= 70 ~ "51-70", 
                                      TRUE ~ "71+")), 
         crimes_per_1000d = fct_relevel(crimes_per_1000d, levels = c("0-23", "24-37", "38-50", "51-70", "71+")))
```

```{r,eval=FALSE}
save(tract_aggregated_2022, file = 'Data/tract_aggregated_2022.RData')
```

Create map of crimes per 1000 people. 
```{r}
ggplot() +
  geom_sf(data=tract_aggregated_2022, aes(fill=crimes_per_1000d), lwd=0.4, alpha =0.8)+
  theme_classic()+
  scale_fill_natparks_d(name = "Olympic")+
  labs(title = "Crimes per 1000 people by Minneapolis census tract 2022", fill = "")+
  theme(plot.title.position = "plot", 
        plot.title = element_text(family = "mono", size = 9), 
        axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank())
```

## Cleaning saint paul crime data

Unlike the Minneapolis crime data, the Saint Paul data did not come with a latitude and longitude location of each crime. However, it did come with a rough block address or street intersection. Using GIS Pro we were able to get a rough latitude and longitude of each crime and then follow the same process as the Minneapolis data.

### Preparing the data for GIS 

First, identify crimes that occurred at intersections. These will need to be treated differently than those with addresses. Use string techniques to put them in a format that GIS likes. 
```{r}
intersections <- saintpaulcrime %>%
  select(BLOCK) %>%
  filter(str_detect(BLOCK, "&")) %>%
  count(BLOCK) %>%
  select(-n)

streets <- as.data.frame(str_split_fixed(intersections$BLOCK, ' & ', 2))
street_identifiers <- as.data.frame(str_split_fixed(streets$V1, " ", 2)) %>%
  #mutate(V2 = word(V2, 1)) %>% trying this out
  rename("street" = "V1", 
         "identifier" = "V2")
  
intersections_RamseyCo <- as.data.frame(str_split_fixed(intersections$BLOCK, ' & ', 2)) %>%
  left_join(street_identifiers, by = c("V2" = "street")) %>%
  mutate(V2 = paste(V2, identifier, sep= " ")) %>%
  select(-identifier) %>%
  rename("street1" = "V1", 
         "street2" = "V2") %>%
  mutate(street1 = str_trim(street1, side = c("right")), 
         street2 = str_trim(street2, side = "right")) %>%
  distinct_all() %>%
  mutate(combined = paste(street1, street2, sep = " & ")) %>%
  select(combined)
```

```{r, eval=FALSE}
write_csv(intersections_RamseyCo, "Data/Crime/intersections_RamseyCo.csv")
```

### Cleaning up GIS data 

Import data and select rough intersection and coordinates.
```{r}
intersections_ll <- read_csv("Data/Crime/intersections_RamseyCo_points.csv")
matched_lat_lngs <- intersections_ll %>%
  filter(Score > 0) %>%
  select(USER_combined, X, Y)
```

Create same intersection descriptions and make column `original` so we are able to join back to original data and get crime date and type. 
```{r}
all_intersections <- saintpaulcrime %>%
  filter(str_detect(BLOCK, "&")) %>%
  select(DATE, INCIDENT, BLOCK)

streets <- as.data.frame(str_split_fixed(all_intersections$BLOCK, ' & ', 2)) %>%
streets2 <- bind_cols(streets, all_intersections)
street_identifiers <- as.data.frame(str_split_fixed(streets2$V1, " ", 2)) %>%
  rename("street" = "V1", 
         "identifier" = "V2")
  
intersections_clean <- streets2 %>%
  left_join(street_identifiers, by = c("V2" = "street")) %>%
  distinct() %>%
  mutate(V2 = paste(V2, identifier, sep= " ")) %>%
  select(-identifier) %>%
  rename("street1" = "V1", 
         "street2" = "V2") %>%
  mutate(street1 = str_trim(street1, side = c("right")), 
         street2 = str_trim(street2, side = "right")) %>%
  mutate(combined = paste(street1, street2, sep = " & ")) 
```

```{r}
intersections_clean %>%
  left_join(matched_lat_lngs, by = c("combined" = "USER_combined")) %>%
  distinct(DATE, X, Y, INCIDENT)
```

### Estimating addresses when given the block

```{r}
saintpaulcrime %>%
  filter(!str_detect(BLOCK, "&")) %>%
  select(DATE, INCIDENT, BLOCK) %>%
  mutate(BLOCK = str_replace(BLOCK, "X", "0")) %>%
  head(100) %>%
  tidygeocoder::geocode(BLOCK, method = 'osm', lat = latitude , long = longitude)
```

