---
title: "Defining Gentrification"
author: "Erin Franke"
date: "9/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the 2nd file in my data cleaning and exploration process. \
Steps in this file: Identify gentrifiable census tracts in 2010. \

## Load libraries and data 

```{r}
library(tidyverse)
library(survey)
library(srvyr)
library(sf)
library(PrettyCols)
library(ggtext)
```

```{r}
roads <- st_read("ShpData/tl_2019_27_prisecroads")
load("Data/acs2010.RData")
load("Data/dicennial2010.RData")
```


## Create organized table of 2010 census tract information

```{r}
tract2010 <- dicennial2010 %>%
  left_join(acs2010, by = c("tract", "county", "GEOID"))
```

## Methods for defining gentrifiable tracts

Lei Ding (2016) of the Federal Reserve Bank of Philadelphia (2016) defines a gentrifiable tract as one that has: 
- median household income is less than the citywide median income

```{r}
gentDing <- tract2010 %>%
  filter(county %in% c("Hennepin", "Ramsey")) %>%
  group_by(county) %>%
  mutate(countyAvgInc = weighted.mean(medianIncome, TotalPopRace), 
         Ding = case_when(medianIncome > countyAvgInc ~ 0, 
                          TRUE ~ 1)) %>%
  select(-countyAvgInc)
```

Professor Lance Freeman (2005): all three measurements below must be met:
- restricts analysis to neighborhoods within the central city of any given metropolitan area. \
- isolates low-income census tracts, using the metropolitan area median housing income as the comparison rather than the citywide median. \
- looks at the share of the housing stock built within the previous 20 years as a measure of relative disinvestment. Tracts in which the share of housing built in the previous 20 years is less than the median for all tracts throughout the metropolitan area are identified.

```{r}
# find metro area mean income
tract2010 %>%
  st_drop_geometry() %>%
  summarize(meanMetroInc = weighted.mean(medianIncome, TotalPopRace, na.rm = TRUE))

# find average amount of housing built in the last twenty years and identify gentrifiable tracts accordingly
gentFreeman <- tract2010 %>%
  filter(county %in% c("Hennepin", "Ramsey")) %>%
  mutate(built1980to2004 = (built2000to2004 + built1990to1999 + built1980to1989)  / totalStructure, 
         medianRecHome = weighted.mean(built1980to2004, totalStructure, na.rm = TRUE), 
         freeman = case_when(medianIncome <= 35167.87 & built1980to2004 <= medianRecHome ~ 1, 
                             TRUE ~ 0)) %>%
  select(freeman, tract) %>%
  st_drop_geometry()
```

Professor Lisa Bates (2013): census tracts must meet at least 3/4 of the criteria \
- a higher percentage of renters than the citywide rate \
- a higher percentage of people of color than citywide \
- a larger share of low-income households than the citywide rate \
- a lower rate of residents with bachelor’s degrees than the citywide rate \

```{r}
gentBates <- tract2010 %>%
  filter(county %in% c("Hennepin", "Ramsey")) %>%
  group_by(county) %>%
  mutate(CityAvgRent = weighted.mean(RentOccupied/OwnershipTotal, OwnershipTotal),
         CityPoc = weighted.mean((TotalPopRace - PopWhiteAlone)/TotalPopRace, TotalPopRace), 
         CityAvgInc = weighted.mean(medianIncome, TotalPopRace), 
         CityBachelors = weighted.mean(bachelors25to64/total25to64, total25to64)) %>%
  ungroup() %>%
  mutate(AvgRent = RentOccupied/OwnershipTotal,
         Poc = (TotalPopRace - PopWhiteAlone)/TotalPopRace, 
         Bachelors = bachelors25to64/total25to64, 
         highRent = case_when(AvgRent >= CityAvgRent ~ 1, 
                                        TRUE ~ 0), 
         highPoc = case_when(Poc >= CityPoc ~ 1, 
                                       TRUE ~ 0), 
         lowInc = case_when(medianIncome <= CityAvgInc ~ 1, 
                                      TRUE ~ 0), 
         lowBachelors = case_when(Bachelors <= CityBachelors ~ 1, 
                                            TRUE ~ 0), 
         totGent = highRent + highPoc + lowInc + lowBachelors, 
         bates = case_when(totGent >= 3 ~ 1, 
                                     TRUE ~0)) %>%
  select(bates, tract) %>%
  st_drop_geometry()
```

## Combine methods and identify gentrifiable tracts

If at least 2/3 of the Bates, Ding, and Freeman methods are met, identify the tract as "gentrifiable" in 2010. 
```{r}
gentrifiable <- gentDing %>%
  left_join(gentFreeman, by = "tract") %>%
  left_join(gentBates, by = "tract") %>%
  mutate(canGent = as.factor(case_when(Ding+freeman + bates >=2 ~ 1, 
                                  TRUE ~ 0))) %>%
  filter(tract != "9800") %>%
  mutate(tract = as.numeric(tract))
```

Create a map of the 2010 census tracts, colored by whether or not the tract can gentrify. 
```{r}
roads_sub <- st_crop(roads,st_bbox(gentrifiable))

gentrifiable %>%
  mutate(canGent = case_when(canGent == "1" ~ "Yes", 
                             TRUE ~ "No")) %>%
  ggplot()+
  geom_sf(aes(fill=canGent), lwd = 0.2, color = "white")+
  geom_sf(data = roads_sub, fill = "gray",color = "gray", lwd= 0.5)+
  theme_classic()+
  labs(title = "Gentrifiable census tracts in Minneapolis-Saint Paul in 2010", fill = "Gentrifiable", x = "", y = "")+
  annotate("point", x = -93.1691, y = 44.938, colour = "orange", size = 1) +
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.title = element_markdown(family = "mono"),
        plot.title.position = "plot", 
        plot.title = element_markdown(family = "mono", size = 10))+
  scale_fill_pretty_d(name = "Autumn", direction = -1)
```

Write the gentrifiable file to the Data folder for future use. 
```{r,eval=FALSE}
save(gentrifiable, file = 'Data/gentrifiable.RData')
```





