---
title: "Determining Gentrified Tracts"
author: "Erin Franke"
date: "2022-09-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the 4th file in my data cleaning and exploration process. \
Steps in this file: Identify census tracts that did and did not gentrify.  \

```{r}
library(tidycensus)
library(tidyverse)
library(survey)
library(srvyr)
library(sf)
library(ggthemes)
load("Data/tractData/cleaned/acs2010clean.RData")
load("Data/tractData/cleaned/acs2020clean.RData")
load("Data/tractData/cleaned/final_condensed.RData")
load("Data/tractData/cleaned/final_condensed_msp.RData")
load("Data/tractData/cleaned/gentrifiable.RData")
roads <- st_read("ShpData/tl_2019_27_prisecroads")
```

## Ding's method for identifying census tracts that gentrified

The tract must see a:
- change in the share of adults with college degrees greater than the city-level change \
- change in median rents above the citywide change OR \
- change in median home value greater than the citywide change. \

```{r}
#citywide changes needed for Ding's method
final_condensed_msp %>%
  group_by(county_2020) %>%
  summarize(avg_propbachelors2010 = weighted.mean(bachelors_2010/population_2010, population_2010), 
         avg_propbachelors2020 = weighted.mean(bachelors_2020/population_2020, population_2020),
         avg_change_bachelors = mean(avg_propbachelors2020 - avg_propbachelors2010),
         avg_home_value_increase = mean((medianHomeValue_2020 - medianHomeValue_2010)/medianHomeValue_2010, na.rm = TRUE),
         avg_percent_rent_increase = mean((medContractRent_2020 - medContractRent_2010)/medContractRent_2010, na.rm = TRUE))
```
  
Using these citywide values, denote each tract as having gentrified or not

```{r}
gent_clean <- gentrifiable %>%
  mutate(propbachelors2010 = bachelors_2010/population_2010, 
         propbachelors2020 = bachelors_2020/population_2020, 
         change_bachelors = propbachelors2020 - propbachelors2010,
         percent_rent_increase = (medContractRent_2020 - medContractRent_2010)/(medContractRent_2010),
         percent_homevalue_increase = (medianHomeValue_2020 - medianHomeValue_2010)/medianHomeValue_2010, 
         gentrified = as.factor(case_when((canGent2010 == "Yes" & city == "Minneapolis" & change_bachelors > 0.03941054 & 
                                       (percent_rent_increase > 0.3259020 | percent_homevalue_increase >0.16874041)) |
                                      (canGent2010 == "Yes" & city == "Saint Paul" & change_bachelors > 0.01854539 &
                                         (percent_rent_increase > 0.3811 | percent_homevalue_increase > 0.05902)) ~ 1, 
                          TRUE ~ 0)), 
         canGent_gentrified = as.factor(case_when(canGent2010 == "No" ~ "Cannot gentrify", 
                                                  canGent2010 == "Yes" & gentrified == "0" ~ "Could, but didn't gentrify", 
                                                  canGent2010 == "Yes" & gentrified == "1" ~ "Gentrified")))
```

Create map of gentrified tracts

```{r}
roads_sub <- st_crop(roads,st_bbox(gentrifiable))

gent_clean %>%
  ggplot()+
  geom_sf(aes(fill=canGent_gentrified), lwd = 0.4)+
  geom_sf(data = roads_sub, fill = "gray",color = "gray", lwd= 0.5)+
  theme_classic()+
  labs(title = "Minneapolis-Saint Paul 2020 census tract gentrification status", fill = "Status", x = "", y = "")+
  #annotate("point", x = -93.1691, y = 44.938, colour = "white", size = 1) +
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.title = element_markdown(family = "mono", size =9),
        plot.title.position = "plot",
        legend.text = element_markdown(family = "mono", size = 8),
        plot.title = element_markdown(family = "mono", size = 10))+
  scale_fill_manual(values = c("#898A99", "#C6431F", "#FFC050"))+
  #annotate(geom = "text", x = -93.1691, y = 44.945, label = "Macalester \nCollege", size = 2, color = "white")+
  guides(color = guide_legend(override.aes = list(size=2,alpha=1)))
```

Understand breakdown of gentrifying census tracts

```{r}
gent_clean %>%
  group_by(city, canGent_gentrified) %>%
  count() %>%
  group_by(city) %>%
  mutate(proportion = n/sum(n))
```

Understand census tracts by variables included in gentrification definition.

```{r}
gent_clean %>%
  mutate(percent_bach = bachelors_2010/population_2010) %>%
  ggplot()+
  geom_sf(aes(fill=percent_bach, color = city), lwd = 0.4)+
  geom_sf(data = roads_sub, fill = "gray",color = "gray", lwd= 0.5)+
  theme_classic()+
  scale_color_manual(values = c("goldenrod", "forestgreen"))+
  labs(title = "MSP census tract gentrification status in 2020", fill = "Gentrifiable", x = "", y = "", color = "City")+
  annotate("point", x = -93.1691, y = 44.938, colour = "white", size = 1) +
  theme(axis.line = element_blank(),
        axis.text = element_blank(), 
        axis.ticks = element_blank(), 
        legend.title = element_markdown(family = "mono"), 
        plot.title.position = "plot", 
        plot.title = element_markdown(family = "mono", size = 12))+
  annotate(geom = "text", x = -93.1691, y = 44.945, label = "Macalester \nCollege", size = 2, color = "white")+
  guides(color = guide_legend(override.aes = list(size=2,alpha=1)))

gent_clean %>%
  mutate(percent_bach = bachelors_2020/population_2020) %>%
  ggplot()+
  geom_sf(aes(fill=percent_bach, color = city), lwd = 0.4)+
  geom_sf(data = roads_sub, fill = "gray",color = "gray", lwd= 0.5)+
  theme_classic()+
  scale_color_manual(values = c("goldenrod", "forestgreen"))+
  labs(title = "Proportion of residents with a bachelors degree by MSP census tract", fill = "Proportion", x = "", y = "", color = "City")+
  annotate("point", x = -93.1691, y = 44.938, colour = "white", size = 1) +
  theme(axis.line = element_blank(),
        axis.text = element_blank(), 
        axis.ticks = element_blank(), 
        legend.title = element_markdown(family = "mono"), 
        plot.title.position = "plot", 
        plot.title = element_markdown(family = "mono", size = 12))+
  annotate(geom = "text", x = -93.1691, y = 44.945, label = "Macalester \nCollege", size = 2, color = "white")+
  guides(color = guide_legend(override.aes = list(size=2,alpha=1)))

gent_clean %>%
  mutate(change_bach = bachelors_2020/population_2020 - bachelors_2010/population_2010) %>%
  ggplot()+
  geom_sf(aes(fill=change_bach, color = city), lwd = 0.4)+
  geom_sf(data = roads_sub, fill = "gray",color = "gray", lwd= 0.5)+
  theme_classic()+
  scale_color_manual(values = c("goldenrod", "forestgreen"))+
  labs(title = "Proportion of residents with a bachelors degree by MSP census tract", fill = "Proportion", x = "", y = "", color = "City")+
  annotate("point", x = -93.1691, y = 44.938, colour = "white", size = 1) +
  scale_fill_gradient2()+
  theme(axis.line = element_blank(),
        axis.text = element_blank(), 
        axis.ticks = element_blank(), 
        legend.title = element_markdown(family = "mono"), 
        plot.title.position = "plot", 
        plot.title = element_markdown(family = "mono", size = 12))+
  annotate(geom = "text", x = -93.1691, y = 44.945, label = "Macalester \nCollege", size = 2, color = "white")+
  guides(color = guide_legend(override.aes = list(size=2,alpha=1)))
```



```{r, eval=FALSE}
save(gent_clean, file = 'Data/tractData/cleaned/gent_clean.RData')
```

