---
title: "Determining Gentrified Tracts"
author: "Erin Franke"
date: "2022-09-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the 4th file in my data cleaning and exploration process. \
Steps in this file: Identify census tracts that did and did not gentrify.  \

```{r}
library(tidycensus)
library(tidyverse)
library(survey)
library(srvyr)
library(sf)
library(ggthemes)
load("Data/tractData/cleaned/acs2010clean.RData")
load("Data/tractData/cleaned/acs2020clean.RData")
load("Data/tractData/cleaned/final_condensed.RData")
load("Data/tractData/cleaned/final_condensed_msp.RData")
load("Data/tractData/cleaned/gentrifiable.RData")
roads <- st_read("ShpData/tl_2019_27_prisecroads")
```

## Methods for idenifying tracts that did gentrify

The Twin Cities paper identifies three research methods of defining gentrification.

Freeman: both required \
- change in the share of adults with college degrees is greater than the regional change \
- tract experienced an increase in home values (in constant dollars).

```{r}
#regional change in share of adults with college degrees
final_condensed %>%
  ungroup() %>%
  summarize(avg_propbachelors2010 = weighted.mean(bachelors25to64_2010/population_2010, population_2010), 
         avg_propbachelors2020 = weighted.mean(bachelors25to64_2020/population_2020, population_2020), 
         avg_change_bachelors = mean(avg_propbachelors2020 - avg_propbachelors2010, na.rm = TRUE))
```

Ding: \
- change in the share of adults with college degrees greater than the city-level change \
- change in median rents above the citywide change OR \
- change in median home value greater than the citywide change. \

```{r}
#citywide changes needed for Ding's method
final_condensed_msp %>%
  group_by(county_2020) %>%
  summarize(avg_propbachelors2010 = weighted.mean(bachelors25to64_2010/population_2010, population_2010), 
         avg_propbachelors2020 = weighted.mean(bachelors25to64_2020/population_2020, population_2020),
         avg_change_bachelors = mean(avg_propbachelors2020 - avg_propbachelors2010),
         avg_home_value_increase = mean((medianHomeValue_2020 - medianHomeValue_2010)/medianHomeValue_2010, na.rm = TRUE),
         avg_percent_rent_increase = mean((medContractRent_2020 - medContractRent_2010)/medContractRent_2010, na.rm = TRUE))
```

Bates methodology was too complex for the data provided and the purpose of this project, so for a third method I looked to a report from the PEW charitable trusts. They defined a tract to have gentrified if the tract’s median income had to have increased at least 10 percent in inflation-adjusted dollars from 2000 to 2014 and and its 2014 median household income figure had to exceed the citywide median of $37,460.https://www.pewtrusts.org/-/media/assets/2016/05/philadelphias_changing_neighborhoods.pdf

```{r}
#regional change in median income
final_condensed_msp %>%
  ungroup() %>%
  group_by(county_2020) %>%
  summarize(medianInc = median(medianIncome_2020))
```
  
## Find gentrified tracts 

Second, create a final `gentrified` dataset that lists whether each or not each census tract was gentrifiable and if it did gentrify according to Ding and Freeman's methods.

```{r}
gent_clean <- gentrifiable %>%
  mutate(propbachelors2010 = bachelors25to64_2010/population_2010, 
         propbachelors2020 = bachelors25to64_2020/population_2020, 
         change_bachelors = propbachelors2020 - propbachelors2010,
         home_value_2010_in_2020dollars = medianHomeValue_2010 * 1.19,
         home_value_increase = medianHomeValue_2020 - home_value_2010_in_2020dollars,
         percent_rent_increase = (medContractRent_2020 - medContractRent_2010)/(medContractRent_2010),
         percent_homevalue_increase = (medianHomeValue_2020 - medianHomeValue_2010)/medianHomeValue_2010,
         medianIncome_2020_in_2010dollars = medianIncome_2020 * 0.84,
         medianIncome2020_10perc_increase = medianIncome_2020_in_2010dollars * 1.1) %>%
  mutate(freeman = as.factor(case_when(canGent == "1" & change_bachelors >= 0.0286 & home_value_increase >= 0 ~ 1, TRUE ~ 0)), 
         ding = as.factor(case_when((canGent == "1" & city == "Minneapolis" & change_bachelors > 0.0394 & 
                                       (percent_rent_increase > 0.3259 | percent_homevalue_increase >0.1687)) |
                                      (canGent == "1" & city == "Saint Paul" & change_bachelors > 0.01955693 &
                                         (percent_rent_increase > 0.38078| percent_homevalue_increase > 0.05902)) ~ 1, 
                          TRUE ~ 0)), 
         pew = as.factor(case_when((canGent == "1" & city == "Minneapolis" & medianIncome_2020 > medianIncome2020_10perc_increase & medianIncome_2020 >= 36776) | (canGent == "1" & city == "Saint Paul" & medianIncome_2020 > medianIncome2020_10perc_increase & medianIncome_2020 >= 31951) ~ 1, 
          TRUE ~ 0)), 
         twoGentrified = as.factor(case_when((ding == "1" & freeman == "1" & pew == "1")|
                                   (ding == "1" & freeman == "1" & pew == "0")|
                                      (ding == "1" & freeman == "0" & pew == "1")|
                                   (ding == "0" & freeman == "1" & pew == "1") ~ 1, 
                                 TRUE ~ 0)), 
         canGent_gentrified = as.factor(case_when(canGent == "0" ~ "0", 
                                                  canGent == "1" & twoGentrified == "0" ~ "1", 
                                                  TRUE ~ "2")))
```

Understand the distribution of gentrified tracts for the different methods.

```{r}
gent_clean %>%
  filter(canGent_gentrified != "0") %>%
  count(freeman, ding, pew, twoGentrified)
```

Create map of gentrified tracts

```{r}
roads_sub <- st_crop(roads,st_bbox(gentrifiable))

gent_clean %>%
  mutate(canGent_gentrified = as.factor(case_when(canGent_gentrified == "0" ~ "Cannot gentrify", 
                                        canGent_gentrified == "1" ~ "Could, but didn't gentrify",
                                        TRUE ~ "Gentrified"))) %>%
  ggplot()+
  geom_sf(aes(fill=canGent_gentrified, color = city), lwd = 0.4)+
  geom_sf(data = roads_sub, fill = "gray",color = "gray", lwd= 0.5)+
  theme_classic()+
  scale_color_manual(values = c("goldenrod", "forestgreen"))+
  labs(title = "MSP census tract gentrification status in 2020", fill = "Gentrifiable", x = "", y = "", color = "City")+
  annotate("point", x = -93.1691, y = 44.938, colour = "white", size = 1) +
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.title = element_markdown(family = "mono"),
        plot.title.position = "plot",
        legend.text = element_markdown(family = "mono"),
        plot.title = element_markdown(family = "mono", size = 12))+
  scale_fill_manual(values = c("#1B3A54", "#800000", "#CD0F75"))+
  annotate(geom = "text", x = -93.1691, y = 44.945, label = "Macalester \nCollege", size = 2, color = "white")+
  guides(color = guide_legend(override.aes = list(size=2,alpha=1)))
```

Additional maps for each method:

```{r}
gent_clean %>%
  mutate(canGent_gentrified = as.factor(case_when(canGent == "0" ~ "Cannot gentrify", 
                                        canGent == "1" & ding == "0" ~ "Could, but didn't gentrify",
                                        canGent == "1" & ding == "1" ~ "Gentrified"))) %>%
  ggplot()+
  geom_sf(aes(fill=canGent_gentrified, color = city), lwd = 0.4)+
  geom_sf(data = roads_sub, fill = "gray",color = "gray", lwd= 0.5)+
  theme_classic()+
  scale_color_manual(values = c("goldenrod", "forestgreen"))+
  labs(title = "MSP census tract gentrification status in 2020 as denoted by Lei Ding", fill = "Gentrifiable", x = "", y = "", color = "City")+
  annotate("point", x = -93.1691, y = 44.938, colour = "white", size = 1) +
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.title = element_markdown(family = "mono"),
        plot.title.position = "plot", 
        plot.title = element_markdown(family = "mono", size = 10))+
  #scale_fill_pretty_d(name = "Autumn", direction = -1)+
  scale_fill_manual(values = c("#1B3A54", "#800000", "#CD0F75"))+
  annotate(geom = "text", x = -93.1691, y = 44.945, label = "Macalester \nCollege", size = 2, color = "white")+
  guides(color = guide_legend(override.aes = list(size=2,alpha=1)))

gent_clean %>%
  mutate(canGent_gentrified = as.factor(case_when(canGent == "0" ~ "Cannot gentrify", 
                                        canGent == "1" & freeman == "0" ~ "Could, but didn't gentrify",
                                        canGent == "1" & freeman == "1" ~ "Gentrified"))) %>%
  ggplot()+
  geom_sf(aes(fill=canGent_gentrified, color = city), lwd = 0.4)+
  geom_sf(data = roads_sub, fill = "gray",color = "gray", lwd= 0.5)+
  theme_classic()+
  scale_color_manual(values = c("goldenrod", "forestgreen"))+
  labs(title = "MSP census tract gentrification status in 2020 as denoted by Lance Freeman", fill = "Gentrifiable", x = "", y = "", color = "City")+
  annotate("point", x = -93.1691, y = 44.938, colour = "white", size = 1) +
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.title = element_markdown(family = "mono"),
        plot.title.position = "plot", 
        plot.title = element_markdown(family = "mono", size = 10))+
  #scale_fill_pretty_d(name = "Autumn", direction = -1)+
  scale_fill_manual(values = c("#1B3A54", "#800000", "#CD0F75"))+
  annotate(geom = "text", x = -93.1691, y = 44.945, label = "Macalester \nCollege", size = 2, color = "white")+
  guides(color = guide_legend(override.aes = list(size=2,alpha=1)))

gent_clean %>%
  mutate(canGent_gentrified = as.factor(case_when(canGent == "0" ~ "Cannot gentrify", 
                                        canGent == "1" & pew == "0" ~ "Could, but didn't gentrify",
                                        canGent == "1" & pew == "1" ~ "Gentrified"))) %>%
  ggplot()+
  geom_sf(aes(fill=canGent_gentrified, color = city), lwd = 0.4)+
  geom_sf(data = roads_sub, fill = "gray",color = "gray", lwd= 0.5)+
  theme_classic()+
  scale_color_manual(values = c("goldenrod", "forestgreen"))+
  labs(title = "MSP census tract gentrification status in 2020 as denoted \nby the PEW Charitiable Trusts", fill = "Gentrifiable", x = "", y = "", color = "City")+
  annotate("point", x = -93.1691, y = 44.938, colour = "white", size = 1) +
  theme(axis.line = element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        legend.title = element_markdown(family = "mono"),
        plot.title.position = "plot", 
        plot.title = element_markdown(family = "mono", size = 10))+
  #scale_fill_pretty_d(name = "Autumn", direction = -1)+
  scale_fill_manual(values = c("#1B3A54", "#800000", "#CD0F75"))+
  annotate(geom = "text", x = -93.1691, y = 44.945, label = "Macalester \nCollege", size = 2, color = "white")+
  guides(color = guide_legend(override.aes = list(size=2,alpha=1)))
```


```{r, eval=FALSE}
save(gent_clean, file = 'Data/tractData/cleaned/gent_clean.RData')
```

