---
title: "Determining Gentrified Tracts"
author: "Erin Franke"
date: "2022-09-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the 3rd file in my data cleaning and exploration process. \
Steps in this file: Identify and deal with census tracts that changed from 2010 to 2020; identify census tracts that did and did not gentrify.  \

```{r}
library(tidycensus)
library(tidyverse)
library(survey)
library(srvyr)
library(sf)
load("Data/acs2010.RData")
load("Data/acs2020.RData")
load("Data/gentrifiable.RData")
```

## Identifying census tract changes from 2010 to 2020

The census tract borders are not the same moving from 2010 to 2020. In order determine gentrified tracts, the first step is to find discrepancies.

The following code shows tracts that did not properly join, meaning tract numbers that were added/deleted when moving from 2010 to 2020.
```{r}
acs2010metro <- acs2010 %>%
  filter(county %in% c("Hennepin", "Ramsey")) %>%
  arrange(GEOID) %>%
  mutate(year2010 = 2010) %>%
  select(tract, year2010, county)
acs2020metro <- acs2020  %>%
  filter(county %in% c("Hennepin", "Ramsey")) %>%
  arrange(GEOID) %>%
  mutate(year2020 = 2020) %>%
  select(tract, year2020, county)

acs2020metro %>%
  full_join(acs2010metro, by= c("tract")) %>%
  filter(is.na(year2020) | is.na(year2010) |is.na(county.x)|is.na(county.y)| county.x != county.y) %>% 
  mutate(tract = as.numeric(tract)) %>%
  arrange(tract)
```

Changes are recorded here: 
- Hennepin 38 in 2010 --> 38.01, 38.02 in 2020 \
- Hennepin 202 --> 202.01, 202.02 \
- Hennepin 232 --> 232.01, 232.02 \
- Hennepin 234 --> 234.01, 234.02 \
- Hennepin 249.01, 249.02 --> 249.04 in 2020 (both 249.01 and 249.02 are gentrifiable) \
- Hennepin 257.01 --> 257.03, 257.04 in 2020 \
- Hennepin 259.03 --> 259.08, 259.09 in 2020 \
- Hennepin 260.16 --> 260.23, 260.24 in 2020 \
- Hennepin 260.18 --> 260.25, 260.26 in 2020 \
- Hennepin 260.22 --> 260.27, 260.28 in 2020 \
- Hennepin 264.02 --> 264.05, 264.06 in 2020 \
- Hennepin 265.12 --> 265.15, 265.16 \
- Hennepin 266.12 --> 266.14, 266.15 \
- Hennepin 266.13 --> 266.16, 266.17 \
- Hennepin 267.10 --> 267.24, 267.25, 267.26 \
- Hennepin 267.14 --> 267.17, 267.18 \
- Hennepin 267.15 --> 267.19, 267.22, 267.23 \
- Hennepin 267.16 --> 267.20, 267.21 \
- Hennepin 268.10 --> 268.27, 268.28 \
- Hennepin 268.20 --> 268.24, 268.25, 268.26 \
- Hennepin 269.09	--> 269.11, 269.12 \
- Hennepin 272.01 --> 272.04, 272.05 \
- Hennepin 277.00 --> 277.01, 277.02, 277.03 \
- Ramsey 342.02 --> 342.03, 342.04 \
- Ramsey 376.02 --> 376.03, 376.04 \
- Ramsey 401.00 --> 401.01, 401.02 \
- Ramsey 406.04 --> 406.05, 406.06 \
- Ramsey 407.03 --> Ramsey 407.08, 407.09 \
- Ramsey 408.02, 408.03 --> 408.04, 408.05 (tract borders change) \
- Ramsey 430.00 --> 430.01, 430.02 \
- Hennepin 1023.00 + 1029.00 --> 1263.00 in 2020 (both gentrifiable) \
- Hennepin 1040.00 --> 1040.01, 1040.02 \
- Hennepin 1048.00 --> 1048.01, 1048.02 \
- Hennepin 1049.00 --> 1049.01, 1049.02 \
- Hennepin 1261.00 --> 1261.01, 1261.02 \
- Hennepin 1262.00 --> 1262.01, 1262.02 \

## Clean the data
Now, make changes to the data. First, change column names of ACS datasets to include the year so that when joining it is clear which value belongs to which.
```{r}
colnames(acs2020) <-paste(colnames(acs2020),"2020",sep="_")
colnames(acs2010) <-paste(colnames(acs2010), "2010", sep="_")
```
 
Clean the ACS 2020 data, creating a modified tract column that represents the original 2010 tract number for all tracts splitting into two or three tracts in 2020. Do not include tracts 9800 or 9801, which is was very unclear what happened to over the 10 year period. Tracts 408.04 and 408.05 must be dealt with specially, as these tracts were labeled 408.02 and 408.03 in 2010 but the borders changed over the decade.
```{r}
acs2020_clean <- acs2020 %>%
  mutate(tract_2020 = as.numeric(tract_2020)) %>%
  filter(county_2020 %in% c("Hennepin", "Ramsey")) %>%
  mutate(tractmod = case_when(tract_2020 %in% c(38.01,38.02) ~ 38.00, 
                               tract_2020 %in% c(202.01, 202.02) ~ 202.00, 
                               tract_2020 %in% c(232.01, 232.02) ~ 232.00, 
                               tract_2020 %in% c(234.01, 234.02) ~ 234.00, 
                               tract_2020 %in% c(257.03, 257.04) ~ 257.01, 
                               tract_2020 %in% c(259.08, 259.09) ~ 259.03, 
                               tract_2020 %in% c(260.23, 260.24) ~ 260.16, 
                               tract_2020 %in% c(260.25, 260.26) ~ 260.18, 
                               tract_2020 %in% c(260.27, 260.28) ~ 260.22, 
                               tract_2020 %in% c(264.05, 264.06) ~ 264.02, 
                               tract_2020 %in% c(265.15, 265.16) ~ 265.12, 
                               tract_2020 %in% c(266.14, 266.15) ~ 266.12, 
                               tract_2020 %in% c(266.16, 266.17) ~ 266.13, 
                               tract_2020 %in% c(267.24, 267.25, 267.26) ~ 267.10, 
                               tract_2020 %in% c(267.17, 267.18) ~ 267.14, 
                               tract_2020 %in% c(267.19, 267.22, 267.23) ~ 267.15, 
                               tract_2020 %in% c(267.20, 267.21) ~ 267.16, 
                               tract_2020 %in% c(268.27, 268.28) ~ 268.10, 
                               tract_2020 %in% c(268.24, 268.25, 268.26) ~ 268.20, 
                               tract_2020 %in% c(269.11, 269.12) ~ 269.09, 
                               tract_2020 %in% c(272.04, 272.05) ~ 272.01, 
                               tract_2020 %in% c(277.01, 277.02, 277.03) ~ 277.00, 
                               tract_2020 %in% c(342.03, 342.04) ~ 342.02,
                               tract_2020 %in% c(376.03, 376.04) ~ 376.02, 
                               tract_2020 %in% c(401.01, 401.02) ~ 401.00, 
                               tract_2020 %in% c(406.05, 406.06) ~ 406.04, 
                               tract_2020 %in% c(407.08, 407.09) ~ 407.03, 
                               tract_2020 %in% c(430.01, 430.02) ~ 430.00, 
                               tract_2020 %in% c(1040.01, 1040.02)~ 1040.00,
                               tract_2020 %in% c(1048.01, 1048.02) ~ 1048.00, 
                               tract_2020 %in% c(1049.01, 1049.02) ~ 1049.00, 
                               tract_2020 %in% c(1261.01, 1261.02) ~ 1261.00,
                               tract_2020 %in% c(1262.01, 1262.02) ~ 1262.00)) %>%
  mutate(tractmod = case_when(is.na(tractmod) ~ tract_2020,
                              TRUE ~ tractmod)) %>%
  filter(tract_2020 != 9800, tract_2020 != 9801, 
         tract_2020 != 408.04, tract_2020 != 408.05)

acs2020groups <- acs2020 %>%
  mutate(tract_2020 = as.numeric(tract_2020)) %>%
  filter(tract_2020 %in% c(408.04, 408.05)) %>%
  summarize(medianIncome_2020 = mean(medianIncome_2020),
            aggregateIncome_2020 = mean(aggregateIncome_2020),
            total25to64_2020 = mean(total25to64_2020),
            bachelors25to64_2020 = mean(bachelors25to64_2020),
            medYrStructureBuilt_2020 = mean(medYrStructureBuilt_2020),
            totalStructure_2020 = mean(totalStructure_2020),
            builtSince2005_2020 = mean(builtSince2005_2020),
            built2000to2004_2020 = mean(built2000to2004_2020),
            built1990to1999_2020 = mean(built1990to1999_2020),
            built1980to1989_2020 = mean(built1980to1989_2020),
            built1970to1979_2020 = mean(built1970to1979_2020),
            built1960to1969_2020 = mean(built1960to1969_2020),
            built1950to1959_2020 = mean(built1950to1959_2020),
            built1940to1949_2020 = mean(built1940to1949_2020),
            builtPre1940_2020 = mean(builtPre1940_2020),
            medContractRent_2020 = mean(medContractRent_2020),
            medGrossRent_2020 = mean(medGrossRent_2020), 
            medianHomeValue_2020 = mean(medianHomeValue_2020)) %>%
  summarize(across(where(is.double), round)) %>%
  mutate(tract_2020 = 408.04, tractmod = 408.04, county_2020 = "Hennepin")

acs2020clean <- acs2020_clean %>%
  bind_rows(acs2020groups)
```

Do a similar process for 2010 tracts. There are a few problematic tracts. 249.01 and 249.02 in 2010 became one tract (249.04) in 2020. Tracts 1023.00 and 1029.00 in 2010 became 1263.00 in 2020.
```{r}
acs2010groups <- acs2010 %>%
  mutate(tract_2010 = as.numeric(tract_2010)) %>%
  filter(tract_2010 %in% c(249.01, 249.02, 1023.00, 1029.00, 408.02, 408.03)) %>%
  mutate(grouped = case_when(tract_2010 %in% c(249.01, 249.02) ~ 249.04, 
                             tract_2010 %in% c(1023.00, 1029.00) ~ 1263.00,
                             TRUE ~ 408.04), 
         population = c(1032, 2383, 1347, 1148, 1778, 4774)) %>%
  group_by(grouped) %>%
  summarize(medianIncome_2010 = weighted.mean(medianIncome_2010, population),
            aggregateIncome_2010 = weighted.mean(aggregateIncome_2010, population),
            total25to64_2010 = weighted.mean(total25to64_2010, population),
            bachelors25to64_2010 = weighted.mean(bachelors25to64_2010, population),
            medYrStructureBuilt_2010 = weighted.mean(medYrStructureBuilt_2010, population),
            totalStructure_2010 = weighted.mean(totalStructure_2010, population),
            builtSince2005_2010 = weighted.mean(builtSince2005_2010, population),
            built2000to2004_2010 = weighted.mean(built2000to2004_2010, population),
            built1990to1999_2010 = weighted.mean(built1990to1999_2010, population),
            built1980to1989_2010 = weighted.mean(built1980to1989_2010, population),
            built1970to1979_2010 = weighted.mean(built1970to1979_2010, population),
            built1960to1969_2010 = weighted.mean(built1960to1969_2010, population),
            built1950to1959_2010 = weighted.mean(built1950to1959_2010, population),
            built1940to1949_2010 = weighted.mean(built1940to1949_2010, population),
            builtPre1940_2010 = weighted.mean(builtPre1940_2010, population),
            medContractRent_2010 = weighted.mean(medContractRent_2010, population),
            medGrossRent_2010 = weighted.mean(medGrossRent_2010, population), 
            medianHomeValue_2010 = weighted.mean(medianHomeValue_2010, population)) %>%
  summarize(across(where(is.double), round)) %>%
  mutate(grouped = c(249.04, 408.04, 1263.00)) %>%
  rename("tract_2010" = "grouped") %>%
  mutate(county_2010 = c("Hennepin", "Ramsey", "Hennepin"))

acs2010clean <- acs2010 %>%
  mutate(tract_2010 = as.numeric(tract_2010)) %>%
  filter(!(tract_2010 %in% c(249.01, 249.02, 1023.00, 1029.00, 408.02, 408.03))) %>%
  bind_rows(acs2010groups)
```

Finally, join the 2010 and 2020 data together.
```{r}
final <- acs2020clean %>%
  left_join(acs2010clean, by = c("tractmod" = "tract_2010")) %>%
  select(tractmod, tract_2020, everything())
```

## Methods for idenifying tracts that did gentrify

The Twin Cities paper identifies three research methods of defining gentrification.

Freeman: both required\
- change in the share of adults with college degrees is greater than the regional change \
- tract experienced an increase in home values (in constant dollars).

Ding: \
- change in the share of adults with college degrees greater than the city-level change \
- change in median rents above the citywide change OR \
- change in median home value greater than the citywide change. \

Bates methodology - complex and did not look to much further into this \
- Groups vulnerable tracts into 3 types of housing market changes \
  -  “Adjacent” tract: low to moderate housing values at the beginning of the study period 
  that did not appreciate significantly during the study period (bottom three quintiles of 
  growth) but borders tracts with high housing values. \
  - “Accelerating” tract is one that was lower to moderate value at the beginning of the study 
  period and had high rates of appreciation (top two quintiles of growth) between 2000 and 2015.
  - “Appreciated” tracts are those that had low to moderate home values in 1990 but had 
  appreciated significantly by 2015.
  
## Find gentrified tracts 

First, calculate average levels of change according to both Freeman and Ding's methods
```{r}
# in hennepin and ramsey county (separate)
final %>%
  group_by(county_2020) %>%
  summarize(avg_propbachelors2010 = mean(bachelors25to64_2010/total25to64_2010, na.rm = TRUE), 
         avg_propbachelors2020 = mean(bachelors25to64_2020/total25to64_2020), 
         avg_change_bachelors = mean(avg_propbachelors2020 - avg_propbachelors2010, na.rm = TRUE),
         avg_home_value_increase = mean(medianHomeValue_2020 - medianHomeValue_2010, na.rm = TRUE),
         avg_percent_rent_increase = mean((medContractRent_2020 - medContractRent_2010)/medContractRent_2010, na.rm =TRUE))

#regional bachelors increase
final %>%
  summarize(avg_propbachelors2010 = mean(bachelors25to64_2010/total25to64_2010, na.rm = TRUE), 
         avg_propbachelors2020 = mean(bachelors25to64_2020/total25to64_2020), 
         avg_change_bachelors = mean(avg_propbachelors2020 - avg_propbachelors2010, na.rm = TRUE), )
```

Second, create a final `gentrified` dataset that lists whether each or not each census tract was gentrifiable and if it did gentrify according to Ding and Freeman's methods.
```{r}
gent_clean <- gentrifiable %>%
  mutate(group = case_when(tract %in% c(249.01, 249.02) ~ "a", 
                           tract %in% c(1023.00, 1029.00) ~ "b",
                           tract %in% c(408.02, 408.03) ~ "c", 
                           TRUE ~ as.character(tract))) %>%
  group_by(group) %>% 
  mutate(geometry = st_union(geometry)) %>% #need to fix tract names (make 249.01 and 249.02 be 249.04)
  
  select(tract, county, canGent) %>%
  mutate(tract = case_when(tract %in% c(249.01, 249.02) ~ 249.04, 
                           tract %in% c(1023.00, 1029.00) ~ 1263.00, 
                           tract %in% c(408.02, 408.03) ~ 408.04, 
                           TRUE ~ tract)) %>%
  distinct_all() %>%
  left_join(final, by = c("tract" = "tractmod")) %>% 
  group_by(tract, county, canGent) %>%
  summarize(bachelors25to64_2010 = mean(bachelors25to64_2010), 
            total25to64_2010 = mean(total25to64_2010), 
            bachelors25to64_2020 = mean(bachelors25to64_2020), 
            total25to64_2020 = mean(total25to64_2020), 
            medianHomeValue_2020 = mean(medianHomeValue_2020), 
            medianHomeValue_2010 = mean(medianHomeValue_2010), 
            medContractRent_2020 = mean(medContractRent_2020), 
            medContractRent_2010 = mean(medContractRent_2010)) %>%
  group_by(tract, canGent, county) %>%
  summarize(propbachelors2010 = bachelors25to64_2010/total25to64_2010, 
         propbachelors2020 = bachelors25to64_2020/total25to64_2020, 
         change_bachelors = propbachelors2020 - propbachelors2010,
         home_value_increase = medianHomeValue_2020 - medianHomeValue_2010,
         percent_rent_increase = (medContractRent_2020 - medContractRent_2010)/(medContractRent_2010)) %>%
  mutate(freeman = as.factor(case_when(canGent == "1" & change_bachelors > 0.06020412 & home_value_increase > 0 ~ 1, TRUE ~ 0)), 
         ding = as.factor(case_when((canGent == "1" & county == "Hennepin" & change_bachelors > 0.06762064 & 
                                       (percent_rent_increase > 0.3581635 | percent_rent_increase > 37945.32)) |
                                      (canGent == "1" & county == "Ramsey" & change_bachelors > 0.04282828 &
                                         (percent_rent_increase > 0.3749619| percent_rent_increase > 18120.00)) ~ 1, 
                          TRUE ~ 0)))
```

Understand what proportion of tracts that could gentrify did. Out of the 216 tracts, Freeman and Ding agreed that 122 did not gentrify and 43 did. Freeman's methods yielded 44 gentrifying that Ding did not, while Ding's methods identified 7 gentrifying the Freeman did not.
```{r}
gent_clean %>%
  filter(canGent == "1") %>%
  ungroup()%>%
  count(freeman, ding)
```

