---
title: "Honors Data Exploration"
author: "Erin Franke"
date: "7/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is my first file in my data cleaning and exploration process. \
Steps in this file: import data using Tidycensus and choose variables of interest. \

## Load libraries 

```{r}
library(tidycensus)
library(tidyverse)
library(survey)
library(srvyr)
library(sf)
```

## Import Data using tidycensus

```{r}
census_api_key("f0d9e7b9eb0a91da1b9d33277cf20e264c19e197", install = TRUE, overwrite=TRUE)
```

The beginning of this study is 2010. Understand what variables are available through the American Community Survey in 2010 (which represents the 2006-2010). 

```{r}
# View variables from 2006 - 2010 ACS (for additional income/education information)
acsvariables2010 <- load_variables(year = 2010, dataset = "acs5") 

# Select variables of interest from 2010 ACS
acs2010 <- get_acs(year = 2010, state = "MN", geography = "tract", variables = c(	
"B01003_001", "B06011_001", "B19313_001", "B23006_001", "B23006_023", "B25035_001", "B25034_001", "B25034_002", "B25034_003", "B25034_004", "B25034_005", "B25034_006", "B25034_007", "B25034_008", "B25034_009", "B25034_010", "B25058_001", "B25064_001", "B25107_001", "B25003_001", "B25003_002", "B25003_003", "B02001_001", "B02001_002"), output = 'wide') %>%
  filter(word(NAME, 4) %in% c("Hennepin", "Ramsey"))%>%
               mutate(tract = word(NAME, 3),
                      tract = str_remove(tract, ","),
                      county = word(NAME, 4)) %>%
               select(-NAME) %>%
               rename("population" = "B01003_001E", 
                      "medianIncome" = "B06011_001E", 
                      "aggregateIncome" = "B19313_001E", 
                      "bachelors25to64" = "B23006_023E",
                      "total25to64" = "B23006_001E", 
                      "medYrStructureBuilt" = "B25035_001E", 
                      "totalStructure" = "B25034_001E", 
                      "builtSince2005" = "B25034_002E", 
                      "built2000to2004" = "B25034_003E", 
                      "built1990to1999" = "B25034_004E", 
                      "built1980to1989" = "B25034_005E", 
                      "built1970to1979" = "B25034_006E", 
                      "built1960to1969" = "B25034_007E", 
                      "built1950to1959" = "B25034_008E", 
                      "built1940to1949" = "B25034_009E", 
                      "builtPre1940" = "B25034_010E", 
                      "medContractRent" = "B25058_001E", 
                      "medGrossRent" = "B25064_001E",
                      "tenureTotal" = "B25003_001E", 
                      "tenureOwned" = "B25003_002E", 
                      "tenureRented" = "B25003_003E",
                      "totalRace" = "B02001_001E", 
                      "totalWhite" = "B02001_002E",
                      "populationMoe" = "B01003_001M",
                      "medIncMoe" = "B06011_001M", 
                      "aggIncMoe" = "B19313_001M", 
                      "bachelors25to64Moe" = "B23006_023M",
                      "total25to64Moe" = "B23006_001M", 
                      "medYrStructureBuiltMoe" = "B25035_001M", 
                      "totalStructureMoe" = "B25034_001M", 
                      "builtSince2005Moe" = "B25034_002M", 
                      "built2000to2004Moe" = "B25034_003M", 
                      "built1990to1999Moe" = "B25034_004M", 
                      "built1980to1989Moe" = "B25034_005M", 
                      "built1970to1979Moe" = "B25034_006M", 
                      "built1960to1969Moe" = "B25034_007M", 
                      "built1950to1959Moe" = "B25034_008M", 
                      "built1940to1949Moe" = "B25034_009M", 
                      "builtPre1940Moe" = "B25034_010M", 
                      "medContractRentMoe" = "B25058_001M", 
                      "medGrossRentMoe" = "B25064_001M", 
                      "medianHomeValue" = "B25107_001E", 
                      "medianHomevalueMoe" = "B25107_001M", 
                      "tenureTotalMoe" = "B25003_001M", 
                      "tenureOwnedMoe" = "B25003_002M", 
                      "tenureRentedMoe" = "B25003_003M", 
                      "totalRaceMoe" = "B02001_001M", 
                      "totalWhiteMoe" = "B02001_002M")
```

Finally, we can do the same thing for the 2020 ACS. There is a decennial census for 2020 but the variables we selected from 2010 are not yet included in that dataset, so I just stuck with the ACS.

```{r}
# Select variables of interest from 2020 ACS
acs2020 <- get_acs(year = 2020, state = "MN", geography = "tract", variables = c(	
"B01003_001", "B06011_001", "B19313_001", "B23006_001", "B23006_023", "B25058_001", "B25064_001", "B25107_001", "B25003_001", "B25003_002", "B25003_003", "B02001_001", "B02001_002"), output = "wide") %>%
  filter(word(NAME, 4) %in% c("Hennepin", "Ramsey"))%>%
               mutate(tract = word(NAME, 3),
                      tract = str_remove(tract, ","),
                      county = word(NAME, 4)) %>%
               select(-NAME) %>%
               rename("population" = "B01003_001E", 
                      "medianIncome" = "B06011_001E", 
                      "aggregateIncome" = "B19313_001E", 
                      "bachelors25to64" = "B23006_023E",
                      "total25to64" = "B23006_001E", 
                      "medContractRent" = "B25058_001E", 
                      "medGrossRent" = "B25064_001E",
                      "populationMoe" = "B01003_001M",
                      "tenureTotal" = "B25003_001E", 
                      "tenureOwned" = "B25003_002E", 
                      "tenureRented" = "B25003_003E",
                      "totalRace" = "B02001_001E", 
                      "totalWhite" = "B02001_002E",
                      "medIncMoe" = "B06011_001M", 
                      "aggIncMoe" = "B19313_001M", 
                      "bachelors25to64Moe" = "B23006_023M",
                      "total25to64Moe" = "B23006_001M", 
                      "medContractRentMoe" = "B25058_001M", 
                      "medGrossRentMoe" = "B25064_001M", 
                      "medianHomeValue" = "B25107_001E", 
                      "medianHomevalueMoe" = "B25107_001M",
                      "tenureTotalMoe" = "B25003_001M", 
                      "tenureOwnedMoe" = "B25003_002M", 
                      "tenureRentedMoe" = "B25003_003M", 
                      "totalRaceMoe" = "B02001_001M", 
                      "totalWhiteMoe" = "B02001_002M")
```

Write ACS and decennial files to Data folder.
```{r, eval=FALSE}
save(acs2010, file = 'Data/tractData/originalTidycensus/acs2010.RData')
save(dicennial2010, file = "Data/tractData/originalTidycensus/dicennial2010.RData")
save(acs2020, file = "Data/tractData/originalTidycensus/acs2020.RData")
```




